### 2009-08-02 doc/latex/class_cowichan_t_b_b.tex

\hypertarget{class_cowichan_t_b_b}{
\section{CowichanTBB Class Reference}
\label{class_cowichan_t_b_b}\index{CowichanTBB@{CowichanTBB}}
}
Thread Building Blocks (TBB) implementation.

{\tt \#include $<$cowichan\_\-tbb.hpp$>$}

Inheritance diagram for CowichanTBB::\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[height=2cm]{class_cowichan_t_b_b}
\end{center}
\end{figure}
\subsection*{Protected Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{mandel} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{randmat} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{half} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixIn, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{invperc} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{thresh} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{life} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixIn, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{winnow} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{class_point}{PointVector} points)
\item 
void \hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{norm} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{hull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{outer} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector)
\item 
void \hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{gauss} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{sor} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{product} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} candidate, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{vecdiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)
\end{CompactItemize}

\subsection{Detailed Description}
Thread Building Blocks (TBB) implementation.

Tags: shared memory, data parallel, task based.

\subsection{Member Function Documentation}
\hypertarget{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{
\index{CowichanTBB@{CowichanTBB}!gauss@{gauss}}
\index{gauss@{gauss}!CowichanTBB@{CowichanTBB}}
\subsubsection[{gauss}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::gauss ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}

For description see \hyperlink{index_gauss_sec}{11. Gaussian Elimination} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan}.\hypertarget{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{
\index{CowichanTBB@{CowichanTBB}!half@{half}}
\index{half@{half}!CowichanTBB@{CowichanTBB}}
\subsubsection[{half}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::half ({\bf IntMatrix} {\em matrixIn}, \/  {\bf IntMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}

For description see \hyperlink{index_half_sec}{3. Two-Dimensional Shuffle} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]matrix to shuffle. \item[{\em matrixOut}]shuffled matrix. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan}.\hypertarget{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{
\index{CowichanTBB@{CowichanTBB}!hull@{hull}}
\index{hull@{hull}!CowichanTBB@{CowichanTBB}}
\subsubsection[{hull}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::hull ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}

For description see \hyperlink{index_hull_sec}{9. Convex Hull}

Runs quickhull algorithm until all points have been used up from the original vector. At each step the hull points are marked as used and a new convex hull is computed on the rest of points. The points that have been used up are in the range (n - hn, n), i.e. at the end of pointsIn vector. NOTE: pointsIn vector gets modified by the algorithm.

Implements \hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan}.\hypertarget{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{
\index{CowichanTBB@{CowichanTBB}!invperc@{invperc}}
\index{invperc@{invperc}!CowichanTBB@{CowichanTBB}}
\subsubsection[{invperc}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::invperc ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}

For description see \hyperlink{index_invperc_sec}{4. Invasion Percolation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]filling resistance matrix. \item[{\em mask}]filled cells. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan}.\hypertarget{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{
\index{CowichanTBB@{CowichanTBB}!life@{life}}
\index{life@{life}!CowichanTBB@{CowichanTBB}}
\subsubsection[{life}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::life ({\bf BoolMatrix} {\em matrixIn}, \/  {\bf BoolMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}

For description see \hyperlink{index_life_sec}{6. Game of Life} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]initial world. \item[{\em matrixOut}]final world. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan}.\hypertarget{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{
\index{CowichanTBB@{CowichanTBB}!mandel@{mandel}}
\index{mandel@{mandel}!CowichanTBB@{CowichanTBB}}
\subsubsection[{mandel}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::mandel ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}

For description see \hyperlink{index_mandel_sec}{1. Mandelbrot Set Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan}.\hypertarget{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{
\index{CowichanTBB@{CowichanTBB}!norm@{norm}}
\index{norm@{norm}!CowichanTBB@{CowichanTBB}}
\subsubsection[{norm}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::norm ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}

For description see \hyperlink{index_norm_sec}{8. Point Location Normalization} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to normalize. \item[{\em pointsOut}]normalized points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan}.\hypertarget{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{
\index{CowichanTBB@{CowichanTBB}!outer@{outer}}
\index{outer@{outer}!CowichanTBB@{CowichanTBB}}
\subsubsection[{outer}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::outer ({\bf PointVector} {\em points}, \/  {\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em vector})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}

For description see \hyperlink{index_outer_sec}{10. Outer Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]vector of points. \item[{\em matrix}]resulting distance matrix. \item[{\em vector}]resulting distance vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan}.\hypertarget{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{
\index{CowichanTBB@{CowichanTBB}!product@{product}}
\index{product@{product}!CowichanTBB@{CowichanTBB}}
\subsubsection[{product}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::product ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em candidate}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}

For description see \hyperlink{index_product_sec}{13. Matrix-Vector Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em candidate}]vector X in AX = V. \item[{\em solution}]vector V in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan}.\hypertarget{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{
\index{CowichanTBB@{CowichanTBB}!randmat@{randmat}}
\index{randmat@{randmat}!CowichanTBB@{CowichanTBB}}
\subsubsection[{randmat}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::randmat ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}

For description see \hyperlink{index_randmat_sec}{2. Random Number Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan}.\hypertarget{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{
\index{CowichanTBB@{CowichanTBB}!sor@{sor}}
\index{sor@{sor}!CowichanTBB@{CowichanTBB}}
\subsubsection[{sor}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::sor ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}

For description see \hyperlink{index_sor_sec}{12. Successive Over-Relaxation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan}.\hypertarget{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{
\index{CowichanTBB@{CowichanTBB}!thresh@{thresh}}
\index{thresh@{thresh}!CowichanTBB@{CowichanTBB}}
\subsubsection[{thresh}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::thresh ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}

For description see \hyperlink{index_thresh_sec}{5. Histogram Thresholding} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]image. \item[{\em mask}]image after thresholding. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan}.\hypertarget{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{
\index{CowichanTBB@{CowichanTBB}!vecdiff@{vecdiff}}
\index{vecdiff@{vecdiff}!CowichanTBB@{CowichanTBB}}
\subsubsection[{vecdiff}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} CowichanTBB::vecdiff ({\bf Vector} {\em actual}, \/  {\bf Vector} {\em computed})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}

For description see \hyperlink{index_vecdiff_sec}{14. 1-Norm Vector Difference} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em actual}]first vector. \item[{\em computed}]second vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan}.\hypertarget{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{
\index{CowichanTBB@{CowichanTBB}!winnow@{winnow}}
\index{winnow@{winnow}!CowichanTBB@{CowichanTBB}}
\subsubsection[{winnow}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::winnow ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask}, \/  {\bf PointVector} {\em points})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}

For description see \hyperlink{index_winnow_sec}{7. Weighted Point Selection} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]integer matrix. \item[{\em mask}]boolean matrix. \item[{\em points}]evenly selected points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan}.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_8hpp}{cowichan\_\-tbb.hpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2gauss_8cpp}{gauss.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2half_8cpp}{half.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2hull_8cpp}{hull.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2invperc_8cpp}{invperc.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2life_8cpp}{life.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2mandel_8cpp}{mandel.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2norm_8cpp}{norm.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2outer_8cpp}{outer.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2product_8cpp}{product.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2randmat_8cpp}{randmat.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2sor_8cpp}{sor.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2thresh_8cpp}{thresh.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2vecdiff_8cpp}{vecdiff.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2winnow_8cpp}{winnow.cpp}\end{CompactItemize}