### 2009-08-02 doc/latex/class_game_of_life.tex

\hypertarget{class_game_of_life}{
\section{GameOfLife Class Reference}
\label{class_game_of_life}\index{GameOfLife@{GameOfLife}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_game_of_life_7de73dfe2060ec6d0a5bab65cf771f63}{
\textbf{GameOfLife} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} first, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} second, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)}
\label{class_game_of_life_7de73dfe2060ec6d0a5bab65cf771f63}

\item 
\hypertarget{class_game_of_life_3027e63b5ad8a3263023f3590caf1d2e}{
void \textbf{swap} ()}
\label{class_game_of_life_3027e63b5ad8a3263023f3590caf1d2e}

\item 
void \hyperlink{class_game_of_life_dfa4429cd90a9a8a7539ecb5a4617ae8}{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range) const 
\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_game_of_life_df162a0c6062460fa658ae9b9cd457b6}{sumNeighbours} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} x, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} y) const 
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_game_of_life_8bdab545960821d63692c05b875199db}{
\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} \textbf{\_\-first}}
\label{class_game_of_life_8bdab545960821d63692c05b875199db}

\item 
\hypertarget{class_game_of_life_d772ef5614e6b178fb8d653589f1f1f9}{
\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} \textbf{\_\-second}}
\label{class_game_of_life_d772ef5614e6b178fb8d653589f1f1f9}

\item 
\hypertarget{class_game_of_life_5a16e4e2b0d83c5e3f0e84cc3ff4db5f}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nr}}
\label{class_game_of_life_5a16e4e2b0d83c5e3f0e84cc3ff4db5f}

\item 
\hypertarget{class_game_of_life_eda1db7eabff4386e01714cf621b072d}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_game_of_life_eda1db7eabff4386e01714cf621b072d}

\end{CompactItemize}

\subsection{Detailed Description}
This class does the game of life, and facilitates a ping-pong memory model.

\subsection{Member Function Documentation}
\hypertarget{class_game_of_life_dfa4429cd90a9a8a7539ecb5a4617ae8}{
\index{GameOfLife@{GameOfLife}!operator()@{operator()}}
\index{operator()@{operator()}!GameOfLife@{GameOfLife}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void GameOfLife::operator() (const {\bf Range2D} \& {\em range}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_game_of_life_dfa4429cd90a9a8a7539ecb5a4617ae8}

Performs the game of life operation over the given range. \hypertarget{class_game_of_life_df162a0c6062460fa658ae9b9cd457b6}{
\index{GameOfLife@{GameOfLife}!sumNeighbours@{sumNeighbours}}
\index{sumNeighbours@{sumNeighbours}!GameOfLife@{GameOfLife}}
\subsubsection[{sumNeighbours}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} GameOfLife::sumNeighbours ({\bf index\_\-t} {\em x}, \/  {\bf index\_\-t} {\em y}) const\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_game_of_life_df162a0c6062460fa658ae9b9cd457b6}

Calculate number of peers.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/life.cpp\end{CompactItemize}