### 2009-08-02 doc/latex/class_make_dominant.tex

\hypertarget{class_make_dominant}{
\section{MakeDominant Class Reference}
\label{class_make_dominant}\index{MakeDominant@{MakeDominant}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_make_dominant_f6d79453b946ddd294f9c4d4109fbb5c}{
\textbf{MakeDominant} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} value)}
\label{class_make_dominant_f6d79453b946ddd294f9c4d4109fbb5c}

\item 
void \hyperlink{class_make_dominant_79c1af1cd1e60a39f5385409209c6cef}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows) const 
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_make_dominant_73325f3005d8887ca20b10fd2cdbdb0b}{
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \textbf{\_\-matrix}}
\label{class_make_dominant_73325f3005d8887ca20b10fd2cdbdb0b}

\item 
\hypertarget{class_make_dominant_770833546d40345b9408226ffbcc6417}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{n}}
\label{class_make_dominant_770833546d40345b9408226ffbcc6417}

\item 
\hypertarget{class_make_dominant_b800551decfb81771f83ab869b459cd8}{
const \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{value}}
\label{class_make_dominant_b800551decfb81771f83ab869b459cd8}

\end{CompactItemize}

\subsection{Detailed Description}
Makes a given matrix diagonally dominant by modifying its diagonal elements.

\subsection{Member Function Documentation}
\hypertarget{class_make_dominant_79c1af1cd1e60a39f5385409209c6cef}{
\index{MakeDominant@{MakeDominant}!operator()@{operator()}}
\index{operator()@{operator()}!MakeDominant@{MakeDominant}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MakeDominant::operator() (const {\bf Range} \& {\em rows}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_make_dominant_79c1af1cd1e60a39f5385409209c6cef}

Sets diagonal elements to a given constant.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/outer.cpp\end{CompactItemize}