### 2009-08-02 doc/latex/class_max_reducer.tex

\hypertarget{class_max_reducer}{
\section{MaxReducer Class Reference}
\label{class_max_reducer}\index{MaxReducer@{MaxReducer}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{class_max_reducer_15429c52ea4767f8c4ed60530814a924}{MaxReducer} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} image, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\item 
\hypertarget{class_max_reducer_de8cd5f464b4287fab14fdbe7b701452}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{getMaximum} () const }
\label{class_max_reducer_de8cd5f464b4287fab14fdbe7b701452}

\item 
void \hyperlink{class_max_reducer_80d53943773b8909921c53a2f094d078}{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range)
\item 
\hyperlink{class_max_reducer_7c2c70197b7e05eb9ecacf0db0b8b3b4}{MaxReducer} (\hyperlink{class_max_reducer}{MaxReducer} \&other, split)
\item 
void \hyperlink{class_max_reducer_095078ad7a356586f27a79a97fc40ac9}{join} (const \hyperlink{class_max_reducer}{MaxReducer} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_max_reducer_90f7cb1f4a7460c19caf4419a0ae1c06}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-image}}
\label{class_max_reducer_90f7cb1f4a7460c19caf4419a0ae1c06}

\item 
\hypertarget{class_max_reducer_8635ce5b2d81812d4a7ddc051c961886}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{\_\-max}}
\label{class_max_reducer_8635ce5b2d81812d4a7ddc051c961886}

\item 
\hypertarget{class_max_reducer_c640e0798bc41ef929aa907cc94cec48}{
const \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_max_reducer_c640e0798bc41ef929aa907cc94cec48}

\end{CompactItemize}

\subsection{Detailed Description}
Performs the maximum computation.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_max_reducer_15429c52ea4767f8c4ed60530814a924}{
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\subsubsection[{MaxReducer}]{\setlength{\rightskip}{0pt plus 5cm}MaxReducer::MaxReducer ({\bf IntMatrix} {\em image}, \/  {\bf index\_\-t} {\em nc})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_15429c52ea4767f8c4ed60530814a924}

Initialise max with the lowest possible value. \hypertarget{class_max_reducer_7c2c70197b7e05eb9ecacf0db0b8b3b4}{
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\subsubsection[{MaxReducer}]{\setlength{\rightskip}{0pt plus 5cm}MaxReducer::MaxReducer ({\bf MaxReducer} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_7c2c70197b7e05eb9ecacf0db0b8b3b4}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_max_reducer_095078ad7a356586f27a79a97fc40ac9}{
\index{MaxReducer@{MaxReducer}!join@{join}}
\index{join@{join}!MaxReducer@{MaxReducer}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void MaxReducer::join (const {\bf MaxReducer} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_095078ad7a356586f27a79a97fc40ac9}

Joiner (TBB). \hypertarget{class_max_reducer_80d53943773b8909921c53a2f094d078}{
\index{MaxReducer@{MaxReducer}!operator()@{operator()}}
\index{operator()@{operator()}!MaxReducer@{MaxReducer}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MaxReducer::operator() (const {\bf Range2D} \& {\em range})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_80d53943773b8909921c53a2f094d078}

Calculates the maximum value over the given range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/thresh.cpp\end{CompactItemize}