### 2009-08-02 doc/latex/class_maximum_distance.tex

\hypertarget{class_maximum_distance}{
\section{MaximumDistance Class Reference}
\label{class_maximum_distance}\index{MaximumDistance@{MaximumDistance}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_maximum_distance_53a46dbac3acecf2223f4966cd73cd69}{
\textbf{MaximumDistance} (const \hyperlink{class_point}{PointVector} \&points, const \hyperlink{class_point}{Point} \&p1, const \hyperlink{class_point}{Point} \&p2)}
\label{class_maximum_distance_53a46dbac3acecf2223f4966cd73cd69}

\item 
\hyperlink{class_point}{Point} $\ast$ \hyperlink{class_maximum_distance_8e1dae43c6ec3865ea79dd7caf27ad8f}{getPoint} () const 
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_maximum_distance_1a187dd3950b791434bb5223d6b16882}{getDistance} () const 
\item 
void \hyperlink{class_maximum_distance_69c66442a93a92b18266ba55909f6c15}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range)
\item 
\hyperlink{class_maximum_distance_1cfb6fc5edbc0495b2fe27304063b68e}{MaximumDistance} (\hyperlink{class_maximum_distance}{MaximumDistance} \&other, split)
\item 
void \hyperlink{class_maximum_distance_8e224531b8c5e9a9f9e506f3e674586b}{join} (const \hyperlink{class_maximum_distance}{MaximumDistance} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_maximum_distance_77b4e1cbd087d7f30f85884c1c9d02a4}{
const \hyperlink{class_point}{PointVector} \& \textbf{points}}
\label{class_maximum_distance_77b4e1cbd087d7f30f85884c1c9d02a4}

\item 
\hypertarget{class_maximum_distance_1ff10f979e3f1e837602e818ea0b8160}{
const \hyperlink{class_point}{Point} \& \textbf{p1}}
\label{class_maximum_distance_1ff10f979e3f1e837602e818ea0b8160}

\item 
\hypertarget{class_maximum_distance_1186a242927b7fc1a33fca0a6eb86006}{
const \hyperlink{class_point}{Point} \& \textbf{p2}}
\label{class_maximum_distance_1186a242927b7fc1a33fca0a6eb86006}

\item 
\hypertarget{class_maximum_distance_1f85a3af711d8c7608fb589d427c519f}{
\hyperlink{class_point}{Point} $\ast$ \textbf{maxPoint}}
\label{class_maximum_distance_1f85a3af711d8c7608fb589d427c519f}

\item 
\hypertarget{class_maximum_distance_b49089f3da83cc1a0e233ad6262ac6cc}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{maxCross}}
\label{class_maximum_distance_b49089f3da83cc1a0e233ad6262ac6cc}

\end{CompactItemize}

\subsection{Detailed Description}
Computes the maximum signed distance from a point to a line, given the line and list of points.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_maximum_distance_1cfb6fc5edbc0495b2fe27304063b68e}{
\index{MaximumDistance@{MaximumDistance}!MaximumDistance@{MaximumDistance}}
\index{MaximumDistance@{MaximumDistance}!MaximumDistance@{MaximumDistance}}
\subsubsection[{MaximumDistance}]{\setlength{\rightskip}{0pt plus 5cm}MaximumDistance::MaximumDistance ({\bf MaximumDistance} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_1cfb6fc5edbc0495b2fe27304063b68e}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_maximum_distance_1a187dd3950b791434bb5223d6b16882}{
\index{MaximumDistance@{MaximumDistance}!getDistance@{getDistance}}
\index{getDistance@{getDistance}!MaximumDistance@{MaximumDistance}}
\subsubsection[{getDistance}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} MaximumDistance::getDistance () const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_1a187dd3950b791434bb5223d6b16882}

Gets the signed distance to that point. \hypertarget{class_maximum_distance_8e1dae43c6ec3865ea79dd7caf27ad8f}{
\index{MaximumDistance@{MaximumDistance}!getPoint@{getPoint}}
\index{getPoint@{getPoint}!MaximumDistance@{MaximumDistance}}
\subsubsection[{getPoint}]{\setlength{\rightskip}{0pt plus 5cm}{\bf Point}$\ast$ MaximumDistance::getPoint () const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_8e1dae43c6ec3865ea79dd7caf27ad8f}

Gets the point with maximum signed distance (if it has already been calculated). \hypertarget{class_maximum_distance_8e224531b8c5e9a9f9e506f3e674586b}{
\index{MaximumDistance@{MaximumDistance}!join@{join}}
\index{join@{join}!MaximumDistance@{MaximumDistance}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void MaximumDistance::join (const {\bf MaximumDistance} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_8e224531b8c5e9a9f9e506f3e674586b}

Joiner (TBB). \hypertarget{class_maximum_distance_69c66442a93a92b18266ba55909f6c15}{
\index{MaximumDistance@{MaximumDistance}!operator()@{operator()}}
\index{operator()@{operator()}!MaximumDistance@{MaximumDistance}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MaximumDistance::operator() (const {\bf Range} \& {\em range})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_69c66442a93a92b18266ba55909f6c15}

Calculates the \hyperlink{class_point}{Point} with maximum signed distance.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/hull.cpp\end{CompactItemize}