### 2009-08-02 doc/latex/class_min_max_reducer.tex

\hypertarget{class_min_max_reducer}{
\section{MinMaxReducer Class Reference}
\label{class_min_max_reducer}\index{MinMaxReducer@{MinMaxReducer}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_min_max_reducer_21bb0b674efa16fe293f02e7ddac5baa}{
\textbf{MinMaxReducer} (\hyperlink{class_point}{PointVector} pointsIn)}
\label{class_min_max_reducer_21bb0b674efa16fe293f02e7ddac5baa}

\item 
\hypertarget{class_min_max_reducer_16c5224355e8fde2e50a8189fca1c2b5}{
\hyperlink{class_point}{Point} \textbf{getMinimum} () const }
\label{class_min_max_reducer_16c5224355e8fde2e50a8189fca1c2b5}

\item 
\hypertarget{class_min_max_reducer_af59d922da8dd11500e068f86e9e140a}{
\hyperlink{class_point}{Point} \textbf{getMaximum} () const }
\label{class_min_max_reducer_af59d922da8dd11500e068f86e9e140a}

\item 
void \hyperlink{class_min_max_reducer_966e6358c637bb3c663a3b6547170fd4}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range)
\item 
\hyperlink{class_min_max_reducer_282eda515485b150389c963f1024ce12}{MinMaxReducer} (\hyperlink{class_min_max_reducer}{MinMaxReducer} \&other, split)
\item 
void \hyperlink{class_min_max_reducer_d002be90fb5c5e273b9f0d5fcb1db988}{join} (const \hyperlink{class_min_max_reducer}{MinMaxReducer} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_min_max_reducer_850a77d5dc5ed32204e265cfc619abc2}{
\hyperlink{class_point}{PointVector} \textbf{pointsIn}}
\label{class_min_max_reducer_850a77d5dc5ed32204e265cfc619abc2}

\item 
\hypertarget{class_min_max_reducer_615c82d310905e78ae116be676357c46}{
\hyperlink{class_point}{Point} \textbf{minPoint}}
\label{class_min_max_reducer_615c82d310905e78ae116be676357c46}

\item 
\hypertarget{class_min_max_reducer_e096e17b34d0bc699a3ed349ed5798d8}{
\hyperlink{class_point}{Point} \textbf{maxPoint}}
\label{class_min_max_reducer_e096e17b34d0bc699a3ed349ed5798d8}

\end{CompactItemize}

\subsection{Detailed Description}
Performs the minimum and maximum computations.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_min_max_reducer_282eda515485b150389c963f1024ce12}{
\index{MinMaxReducer@{MinMaxReducer}!MinMaxReducer@{MinMaxReducer}}
\index{MinMaxReducer@{MinMaxReducer}!MinMaxReducer@{MinMaxReducer}}
\subsubsection[{MinMaxReducer}]{\setlength{\rightskip}{0pt plus 5cm}MinMaxReducer::MinMaxReducer ({\bf MinMaxReducer} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_min_max_reducer_282eda515485b150389c963f1024ce12}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_min_max_reducer_d002be90fb5c5e273b9f0d5fcb1db988}{
\index{MinMaxReducer@{MinMaxReducer}!join@{join}}
\index{join@{join}!MinMaxReducer@{MinMaxReducer}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void MinMaxReducer::join (const {\bf MinMaxReducer} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_min_max_reducer_d002be90fb5c5e273b9f0d5fcb1db988}

Joiner (TBB). \hypertarget{class_min_max_reducer_966e6358c637bb3c663a3b6547170fd4}{
\index{MinMaxReducer@{MinMaxReducer}!operator()@{operator()}}
\index{operator()@{operator()}!MinMaxReducer@{MinMaxReducer}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MinMaxReducer::operator() (const {\bf Range} \& {\em range})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_min_max_reducer_966e6358c637bb3c663a3b6547170fd4}

Calculates the minimum and maximum co-ordinates over the given array range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/norm.cpp\end{CompactItemize}