### 2009-08-02 doc/latex/class_norm_diff.tex

\hypertarget{class_norm_diff}{
\section{NormDiff Class Reference}
\label{class_norm_diff}\index{NormDiff@{NormDiff}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_norm_diff_47bd32e56a26272eaad47a4839448312}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{getNormDiff} () const }
\label{class_norm_diff_47bd32e56a26272eaad47a4839448312}

\item 
\hypertarget{class_norm_diff_2f6ff2bc440755f8ac2350b5185b5423}{
\textbf{NormDiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)}
\label{class_norm_diff_2f6ff2bc440755f8ac2350b5185b5423}

\item 
\hypertarget{class_norm_diff_44066147e706abbfce8e16fc92e0affc}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range)}
\label{class_norm_diff_44066147e706abbfce8e16fc92e0affc}

\item 
\hyperlink{class_norm_diff_edef38012c44577650ccb47a55177eee}{NormDiff} (\hyperlink{class_norm_diff}{NormDiff} \&other, split)
\item 
void \hyperlink{class_norm_diff_8a3ce83486dfa068fbc4d3601a6442b1}{join} (const \hyperlink{class_norm_diff}{NormDiff} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_norm_diff_f4dd514a2fb97d4e4742232b2d6f4930}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-actual}}
\label{class_norm_diff_f4dd514a2fb97d4e4742232b2d6f4930}

\item 
\hypertarget{class_norm_diff_c6815c7be114269afb88e6932a001496}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-computed}}
\label{class_norm_diff_c6815c7be114269afb88e6932a001496}

\item 
\hypertarget{class_norm_diff_99c21620f91d9f917bb0745a284275b3}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{normDiff}}
\label{class_norm_diff_99c21620f91d9f917bb0745a284275b3}

\end{CompactItemize}

\subsection{Detailed Description}
This class takes the 1-norm of the difference between two vectors.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_norm_diff_edef38012c44577650ccb47a55177eee}{
\index{NormDiff@{NormDiff}!NormDiff@{NormDiff}}
\index{NormDiff@{NormDiff}!NormDiff@{NormDiff}}
\subsubsection[{NormDiff}]{\setlength{\rightskip}{0pt plus 5cm}NormDiff::NormDiff ({\bf NormDiff} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_norm_diff_edef38012c44577650ccb47a55177eee}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_norm_diff_8a3ce83486dfa068fbc4d3601a6442b1}{
\index{NormDiff@{NormDiff}!join@{join}}
\index{join@{join}!NormDiff@{NormDiff}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void NormDiff::join (const {\bf NormDiff} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_norm_diff_8a3ce83486dfa068fbc4d3601a6442b1}

Joiner (TBB).

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/vecdiff.cpp\end{CompactItemize}