### 2009-08-02 doc/latex/class_normalizer.tex

\hypertarget{class_normalizer}{
\section{Normalizer Class Reference}
\label{class_normalizer}\index{Normalizer@{Normalizer}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_normalizer_56999f560190503e8f001bdc7ecc2b87}{
\textbf{Normalizer} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} minX, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} minY, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} xfactor, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} yfactor)}
\label{class_normalizer_56999f560190503e8f001bdc7ecc2b87}

\item 
\hypertarget{class_normalizer_f384e3a61919ce96f30b52248c384261}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range) const }
\label{class_normalizer_f384e3a61919ce96f30b52248c384261}

\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_normalizer_02f117411f07fae9fc034885e69f298b}{
\hyperlink{class_point}{PointVector} \textbf{pointsIn}}
\label{class_normalizer_02f117411f07fae9fc034885e69f298b}

\item 
\hypertarget{class_normalizer_5f731c7abc673530d5da15a5b73b03f3}{
\hyperlink{class_point}{PointVector} \textbf{pointsOut}}
\label{class_normalizer_5f731c7abc673530d5da15a5b73b03f3}

\item 
\hypertarget{class_normalizer_d928591add8f6af2fc4943b6a3800242}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{minX}}
\label{class_normalizer_d928591add8f6af2fc4943b6a3800242}

\item 
\hypertarget{class_normalizer_e154cea5c227b01fa743a18c20fc94bc}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{minY}}
\label{class_normalizer_e154cea5c227b01fa743a18c20fc94bc}

\item 
\hypertarget{class_normalizer_e25d5be008bd7b298b0e3c2a7969ed77}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{xfactor}}
\label{class_normalizer_e25d5be008bd7b298b0e3c2a7969ed77}

\item 
\hypertarget{class_normalizer_67a219d5d5e172dfca12b0a22bbe53a6}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{yfactor}}
\label{class_normalizer_67a219d5d5e172dfca12b0a22bbe53a6}

\end{CompactItemize}

\subsection{Detailed Description}
This class performs point normalization -- points are put onto the unit square.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/norm.cpp\end{CompactItemize}