### 2009-08-02 doc/latex/class_perc_point.tex

\hypertarget{class_perc_point}{
\section{PercPoint Class Reference}
\label{class_perc_point}\index{PercPoint@{PercPoint}}
}
Percolation point.

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{PercPoint} (\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point})
\item 
bool \hyperlink{class_perc_point_f798aa389b770208bd172542267ec811}{operator$<$} (const \hyperlink{class_perc_point}{PercPoint} \&other) const 
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{value} () const 
\item 
\hypertarget{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{
\textbf{PercPoint} (\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point})}
\label{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}

\item 
\hypertarget{class_perc_point_f798aa389b770208bd172542267ec811}{
bool \textbf{operator$<$} (const \hyperlink{class_perc_point}{PercPoint} \&other) const }
\label{class_perc_point_f798aa389b770208bd172542267ec811}

\item 
\hypertarget{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{value} () const }
\label{class_perc_point_0099784a4e3728c1f28b546a5b94104a}

\item 
\hypertarget{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{
\textbf{PercPoint} (\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point})}
\label{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}

\item 
\hypertarget{class_perc_point_f798aa389b770208bd172542267ec811}{
bool \textbf{operator$<$} (const \hyperlink{class_perc_point}{PercPoint} \&other) const }
\label{class_perc_point_f798aa389b770208bd172542267ec811}

\item 
\hypertarget{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{value} () const }
\label{class_perc_point_0099784a4e3728c1f28b546a5b94104a}

\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point}
\item 
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \hyperlink{class_perc_point_730c461b6a04bcb883d60df9b297b5d5}{matrix}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_perc_point_dc46d54bf942be6e7774ae522e563f87}{nc}
\end{CompactItemize}

\subsection{Detailed Description}
Percolation point.

These are put into a heap to extract the one with lowest value.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{
\index{PercPoint@{PercPoint}!PercPoint@{PercPoint}}
\index{PercPoint@{PercPoint}!PercPoint@{PercPoint}}
\subsubsection[{PercPoint}]{\setlength{\rightskip}{0pt plus 5cm}PercPoint::PercPoint ({\bf Point} {\em point})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}

Construct a percolation point. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em point}]point. \end{description}
\end{Desc}

\subsection{Member Function Documentation}
\hypertarget{class_perc_point_f798aa389b770208bd172542267ec811}{
\index{PercPoint@{PercPoint}!operator$<$@{operator$<$}}
\index{operator$<$@{operator$<$}!PercPoint@{PercPoint}}
\subsubsection[{operator$<$}]{\setlength{\rightskip}{0pt plus 5cm}bool PercPoint::operator$<$ (const {\bf PercPoint} \& {\em other}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_perc_point_f798aa389b770208bd172542267ec811}

Less than comparison - we want to extract lowest values. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em other}]right hand side percolation point. \end{description}
\end{Desc}
\hypertarget{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{
\index{PercPoint@{PercPoint}!value@{value}}
\index{value@{value}!PercPoint@{PercPoint}}
\subsubsection[{value}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} PercPoint::value () const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_perc_point_0099784a4e3728c1f28b546a5b94104a}

Get the value at this point. \begin{Desc}
\item[Returns:]value from the matrix. \end{Desc}

\subsection{Member Data Documentation}
\hypertarget{class_perc_point_730c461b6a04bcb883d60df9b297b5d5}{
\index{PercPoint@{PercPoint}!matrix@{matrix}}
\index{matrix@{matrix}!PercPoint@{PercPoint}}
\subsubsection[{matrix}]{\setlength{\rightskip}{0pt plus 5cm}{\bf IntMatrix} {\bf PercPoint::matrix}}}
\label{class_perc_point_730c461b6a04bcb883d60df9b297b5d5}

Matrix. \hypertarget{class_perc_point_dc46d54bf942be6e7774ae522e563f87}{
\index{PercPoint@{PercPoint}!nc@{nc}}
\index{nc@{nc}!PercPoint@{PercPoint}}
\subsubsection[{nc}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf PercPoint::nc}}}
\label{class_perc_point_dc46d54bf942be6e7774ae522e563f87}

Number of columns. \hypertarget{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{
\index{PercPoint@{PercPoint}!point@{point}}
\index{point@{point}!PercPoint@{PercPoint}}
\subsubsection[{point}]{\setlength{\rightskip}{0pt plus 5cm}{\bf Point} {\bf PercPoint::point}}}
\label{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}

\hyperlink{class_point}{Point}.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2invperc_8cpp}{invperc.cpp}\item 
cowichan\_\-openmp/invperc.cpp\item 
cowichan\_\-tbb/invperc.cpp\end{CompactItemize}