### 2009-08-02 doc/latex/class_point_distances.tex

\hypertarget{class_point_distances}{
\section{PointDistances Class Reference}
\label{class_point_distances}\index{PointDistances@{PointDistances}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_point_distances_12dba51e28ac2cf5ac0c9317d9ff242e}{
\textbf{PointDistances} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)}
\label{class_point_distances_12dba51e28ac2cf5ac0c9317d9ff242e}

\item 
\hypertarget{class_point_distances_ec6ace35911f7e26d573023f2381c673}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{getMaximum} () const }
\label{class_point_distances_ec6ace35911f7e26d573023f2381c673}

\item 
void \hyperlink{class_point_distances_ab3ec991122054f46c2d5b5f7f91e972}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows)
\item 
\hyperlink{class_point_distances_59396cd3119746888d0601cd78757da3}{PointDistances} (\hyperlink{class_point_distances}{PointDistances} \&other, split)
\item 
void \hyperlink{class_point_distances_609701c6fa66e3f0406db82bc70453d9}{join} (const \hyperlink{class_point_distances}{PointDistances} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_point_distances_a6640fdeda32580a760309f564f8ac9d}{
\hyperlink{class_point}{PointVector} \textbf{\_\-points}}
\label{class_point_distances_a6640fdeda32580a760309f564f8ac9d}

\item 
\hypertarget{class_point_distances_3044927709dad1b9a94784f077d83190}{
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \textbf{\_\-matrix}}
\label{class_point_distances_3044927709dad1b9a94784f077d83190}

\item 
\hypertarget{class_point_distances_3233cd7ba2cdb8a51ceadce064569c7d}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-vector}}
\label{class_point_distances_3233cd7ba2cdb8a51ceadce064569c7d}

\item 
\hypertarget{class_point_distances_633e1a7fb762471b9a4806d52e024fcc}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{n}}
\label{class_point_distances_633e1a7fb762471b9a4806d52e024fcc}

\item 
\hypertarget{class_point_distances_5aa37e9525054b926bb2ce5c6bb515a8}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{\_\-max}}
\label{class_point_distances_5aa37e9525054b926bb2ce5c6bb515a8}

\end{CompactItemize}

\subsection{Detailed Description}
This module turns a vector containing point positions into a dense, symmet- ric, diagonally dominant matrix by calculating the distances between each pair of points. It also constructs a real vector whose values are the distance of each point from the origin. Inputs are:

points: a vector of (x, y) points, where x and y are the point’s position. nelts: the number of points in the vector, and the size of the matrix along each axis.

Its outputs are:

matrix: a real matrix, whose values are filled with inter-point distances. vector: a real vector, whose values are filled with origin-to-point distances.

Each matrix element Mi,j such that i != j is given the value di,j, the Eu- clidean distance between point i and point j. The diagonal values Mi,i are then set to nelts times the maximum off-diagonal value to ensure that the matrix is diagonally dominant. The value of the vector element vi is set to the distance of point i from the origin, which is given by sqrt(xi$^\wedge$2 + yi$^\wedge$2).

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_point_distances_59396cd3119746888d0601cd78757da3}{
\index{PointDistances@{PointDistances}!PointDistances@{PointDistances}}
\index{PointDistances@{PointDistances}!PointDistances@{PointDistances}}
\subsubsection[{PointDistances}]{\setlength{\rightskip}{0pt plus 5cm}PointDistances::PointDistances ({\bf PointDistances} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_point_distances_59396cd3119746888d0601cd78757da3}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_point_distances_609701c6fa66e3f0406db82bc70453d9}{
\index{PointDistances@{PointDistances}!join@{join}}
\index{join@{join}!PointDistances@{PointDistances}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void PointDistances::join (const {\bf PointDistances} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_point_distances_609701c6fa66e3f0406db82bc70453d9}

Joiner (TBB). \hypertarget{class_point_distances_ab3ec991122054f46c2d5b5f7f91e972}{
\index{PointDistances@{PointDistances}!operator()@{operator()}}
\index{operator()@{operator()}!PointDistances@{PointDistances}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void PointDistances::operator() (const {\bf Range} \& {\em rows})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_point_distances_ab3ec991122054f46c2d5b5f7f91e972}

Calculates inter-point distances on the given range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/outer.cpp\end{CompactItemize}