### 2009-08-02 doc/latex/class_random_generator.tex

\hypertarget{class_random_generator}{
\section{RandomGenerator Class Reference}
\label{class_random_generator}\index{RandomGenerator@{RandomGenerator}}
}
Random generator.

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_random_generator_bfa47006c53f82e85cbc77ae5e667747}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows) const 
\item 
\hyperlink{class_random_generator_96274661438a80cfee1e7a33bc59f770}{RandomGenerator} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}{nr}, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}{nc})
\item 
void \hyperlink{class_random_generator_4c5ae77662dfca92b1677e36b124e3c4}{execute} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_caea6ecf8e980861f1d370cb54eaeeda}{next} (\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \&current) const 
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_d0ee441c774a78abd93a50c0b08eea43}{nextK} (\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \&current) const 
\item 
void \hyperlink{class_random_generator_8a8390ed7e511770795cf4204d1f1df0}{initialise} ()
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \hyperlink{class_random_generator_9fb2b8631b95d10abb2f013319d37f3f}{\_\-matrix}
\item 
\hyperlink{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}{IntVector} \hyperlink{class_random_generator_4423c35148a29e6ddd8e108a144db2ac}{state}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_3f733ddf90320aace1be173bdc7381da}{aPrime}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_990ae8409e5290b6e911b609fb61d8c8}{cPrime}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}{nr}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}{nc}
\end{CompactItemize}

\subsection{Detailed Description}
Random generator.

This class generates random integers in a matrix, using a simple linear congruential number generator with a given seed value, i.e. using the recurrence:

X\_\-i+1 = (a$\ast$X\_\-i + c) mod m

The class also provides a parallel implementation with parallel\_\-for.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_random_generator_96274661438a80cfee1e7a33bc59f770}{
\index{RandomGenerator@{RandomGenerator}!RandomGenerator@{RandomGenerator}}
\index{RandomGenerator@{RandomGenerator}!RandomGenerator@{RandomGenerator}}
\subsubsection[{RandomGenerator}]{\setlength{\rightskip}{0pt plus 5cm}RandomGenerator::RandomGenerator ({\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_random_generator_96274661438a80cfee1e7a33bc59f770}

Construct a random number generator. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em nr}]number of matrix rows. \item[{\em nc}]number of matrix columns. \end{description}
\end{Desc}

\subsection{Member Function Documentation}
\hypertarget{class_random_generator_4c5ae77662dfca92b1677e36b124e3c4}{
\index{RandomGenerator@{RandomGenerator}!execute@{execute}}
\index{execute@{execute}!RandomGenerator@{RandomGenerator}}
\subsubsection[{execute}]{\setlength{\rightskip}{0pt plus 5cm}void RandomGenerator::execute ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_random_generator_4c5ae77662dfca92b1677e36b124e3c4}

Run in parallel. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to use. \end{description}
\end{Desc}
\hypertarget{class_random_generator_8a8390ed7e511770795cf4204d1f1df0}{
\index{RandomGenerator@{RandomGenerator}!initialise@{initialise}}
\index{initialise@{initialise}!RandomGenerator@{RandomGenerator}}
\subsubsection[{initialise}]{\setlength{\rightskip}{0pt plus 5cm}void RandomGenerator::initialise ()\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_random_generator_8a8390ed7e511770795cf4204d1f1df0}

Initialises the random number generator to operate on NROWS different rows completely independently. \hypertarget{class_random_generator_caea6ecf8e980861f1d370cb54eaeeda}{
\index{RandomGenerator@{RandomGenerator}!next@{next}}
\index{next@{next}!RandomGenerator@{RandomGenerator}}
\subsubsection[{next}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} RandomGenerator::next ({\bf INT\_\-TYPE} \& {\em current}) const\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_random_generator_caea6ecf8e980861f1d370cb54eaeeda}

Generates the next random number. Convenience method for next(current, a, c); \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em current}]current element. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Next element. \end{Desc}
\hypertarget{class_random_generator_d0ee441c774a78abd93a50c0b08eea43}{
\index{RandomGenerator@{RandomGenerator}!nextK@{nextK}}
\index{nextK@{nextK}!RandomGenerator@{RandomGenerator}}
\subsubsection[{nextK}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} RandomGenerator::nextK ({\bf INT\_\-TYPE} \& {\em current}) const\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_random_generator_d0ee441c774a78abd93a50c0b08eea43}

Generates the k-th next random number, using an identity, where A = a$^\wedge$k mod m and C = c $\ast$ sum\mbox{[}j=0..k-1\mbox{]}(a$^\wedge$j mod m). \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em current}]element. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Next k-th element. \end{Desc}
\hypertarget{class_random_generator_bfa47006c53f82e85cbc77ae5e667747}{
\index{RandomGenerator@{RandomGenerator}!operator()@{operator()}}
\index{operator()@{operator()}!RandomGenerator@{RandomGenerator}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void RandomGenerator::operator() (const {\bf Range} \& {\em rows}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_random_generator_bfa47006c53f82e85cbc77ae5e667747}

Provides NROWS-seperated linear congruential RNG on the specified range of rows. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em rows}]range of rows to work on. \end{description}
\end{Desc}

\subsection{Member Data Documentation}
\hypertarget{class_random_generator_9fb2b8631b95d10abb2f013319d37f3f}{
\index{RandomGenerator@{RandomGenerator}!\_\-matrix@{\_\-matrix}}
\index{\_\-matrix@{\_\-matrix}!RandomGenerator@{RandomGenerator}}
\subsubsection[{\_\-matrix}]{\setlength{\rightskip}{0pt plus 5cm}{\bf IntMatrix} {\bf RandomGenerator::\_\-matrix}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_9fb2b8631b95d10abb2f013319d37f3f}

Matrix to fill. \hypertarget{class_random_generator_3f733ddf90320aace1be173bdc7381da}{
\index{RandomGenerator@{RandomGenerator}!aPrime@{aPrime}}
\index{aPrime@{aPrime}!RandomGenerator@{RandomGenerator}}
\subsubsection[{aPrime}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} {\bf RandomGenerator::aPrime}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_3f733ddf90320aace1be173bdc7381da}

Constant a prime. \hypertarget{class_random_generator_990ae8409e5290b6e911b609fb61d8c8}{
\index{RandomGenerator@{RandomGenerator}!cPrime@{cPrime}}
\index{cPrime@{cPrime}!RandomGenerator@{RandomGenerator}}
\subsubsection[{cPrime}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} {\bf RandomGenerator::cPrime}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_990ae8409e5290b6e911b609fb61d8c8}

Constant c prime. \hypertarget{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}{
\index{RandomGenerator@{RandomGenerator}!nc@{nc}}
\index{nc@{nc}!RandomGenerator@{RandomGenerator}}
\subsubsection[{nc}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf RandomGenerator::nc}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}

Number of columns in the matrix. \hypertarget{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}{
\index{RandomGenerator@{RandomGenerator}!nr@{nr}}
\index{nr@{nr}!RandomGenerator@{RandomGenerator}}
\subsubsection[{nr}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf RandomGenerator::nr}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}

Number of rows in the matrix. \hypertarget{class_random_generator_4423c35148a29e6ddd8e108a144db2ac}{
\index{RandomGenerator@{RandomGenerator}!state@{state}}
\index{state@{state}!RandomGenerator@{RandomGenerator}}
\subsubsection[{state}]{\setlength{\rightskip}{0pt plus 5cm}{\bf IntVector} {\bf RandomGenerator::state}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_4423c35148a29e6ddd8e108a144db2ac}

First element vector.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2randmat_8cpp}{randmat.cpp}\end{CompactItemize}