### 2009-08-02 doc/latex/class_row_elimination.tex

\hypertarget{class_row_elimination}{
\section{RowElimination Class Reference}
\label{class_row_elimination}\index{RowElimination@{RowElimination}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_row_elimination_582416939554bbb8e462e53d6d86f8d6}{
\textbf{RowElimination} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)}
\label{class_row_elimination_582416939554bbb8e462e53d6d86f8d6}

\item 
\hypertarget{class_row_elimination_9037d2255c9363e4ed5315e4457e09c2}{
void \textbf{setI} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} i)}
\label{class_row_elimination_9037d2255c9363e4ed5315e4457e09c2}

\item 
\hypertarget{class_row_elimination_8fa4ceac659d237e6cf693321623e328}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows) const }
\label{class_row_elimination_8fa4ceac659d237e6cf693321623e328}

\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_row_elimination_49344262a09cdf006d4b249e8fddf3d4}{
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \textbf{\_\-matrix}}
\label{class_row_elimination_49344262a09cdf006d4b249e8fddf3d4}

\item 
\hypertarget{class_row_elimination_1f473bc02c05aa508b467e6097e13ddf}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-target}}
\label{class_row_elimination_1f473bc02c05aa508b467e6097e13ddf}

\item 
\hypertarget{class_row_elimination_593558f30cdc67d74b41e11197a56095}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{n}}
\label{class_row_elimination_593558f30cdc67d74b41e11197a56095}

\item 
\hypertarget{class_row_elimination_4ef655247f5284b1694b57f5687d2d40}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{i}}
\label{class_row_elimination_4ef655247f5284b1694b57f5687d2d40}

\item 
\hypertarget{class_row_elimination_5a96a12ab6503193733e9e4c2e545f07}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{column\_\-i}}
\label{class_row_elimination_5a96a12ab6503193733e9e4c2e545f07}

\end{CompactItemize}

\subsection{Detailed Description}
Performs row elimination, i.e. eliminate i-th column in j-th row.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/gauss.cpp\end{CompactItemize}