### 2009-08-02 doc/latex/class_shuffle.tex

\hypertarget{class_shuffle}{
\section{Shuffle Class Reference}
\label{class_shuffle}\index{Shuffle@{Shuffle}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_shuffle_b6e81ca9e45260ac098005caddd5a663}{
\textbf{Shuffle} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} input, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} output, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)}
\label{class_shuffle_b6e81ca9e45260ac098005caddd5a663}

\item 
void \hyperlink{class_shuffle_7922aacb5aae207d2e51ded86b3c1c9d}{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range) const 
\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_shuffle_b0fed7e167e632459538da3d6e810b2e}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-first}}
\label{class_shuffle_b0fed7e167e632459538da3d6e810b2e}

\item 
\hypertarget{class_shuffle_4b6d3c7369a12163f9ad5a080a5b208e}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-second}}
\label{class_shuffle_4b6d3c7369a12163f9ad5a080a5b208e}

\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_shuffle_d8f6038cc7fd4f722d7f00f2a6af9698}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{xBreak}}
\label{class_shuffle_d8f6038cc7fd4f722d7f00f2a6af9698}

\item 
\hypertarget{class_shuffle_a41b51804d77e262d65fc79474879e9c}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{yBreak}}
\label{class_shuffle_a41b51804d77e262d65fc79474879e9c}

\item 
\hypertarget{class_shuffle_d685d43331c6f9a9d0f0775d8e21ebff}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nr}}
\label{class_shuffle_d685d43331c6f9a9d0f0775d8e21ebff}

\item 
\hypertarget{class_shuffle_c7f2688b41a3bc24e7939164a28784e0}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_shuffle_c7f2688b41a3bc24e7939164a28784e0}

\end{CompactItemize}

\subsection{Detailed Description}
This class does a halving shuffle.

\subsection{Member Function Documentation}
\hypertarget{class_shuffle_7922aacb5aae207d2e51ded86b3c1c9d}{
\index{Shuffle@{Shuffle}!operator()@{operator()}}
\index{operator()@{operator()}!Shuffle@{Shuffle}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void Shuffle::operator() (const {\bf Range2D} \& {\em range}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_shuffle_7922aacb5aae207d2e51ded86b3c1c9d}

Performs the halving shuffle over the given range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/half.cpp\end{CompactItemize}