### 2009-08-02 doc/latex/class_threshold.tex

\hypertarget{class_threshold}{
\section{Threshold Class Reference}
\label{class_threshold}\index{Threshold@{Threshold}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_threshold_bdd06b7735ea6b46efab1685411e3d70}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range) const }
\label{class_threshold_bdd06b7735ea6b46efab1685411e3d70}

\item 
\hypertarget{class_threshold_df53dd36b3ab7336627ab7361ab3eb97}{
\textbf{Threshold} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} image, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} retain, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} result, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)}
\label{class_threshold_df53dd36b3ab7336627ab7361ab3eb97}

\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_threshold_53a4f784d8bd2d813a21aa1a52e5afbb}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-image}}
\label{class_threshold_53a4f784d8bd2d813a21aa1a52e5afbb}

\item 
\hypertarget{class_threshold_f5e82e77ffad10da1dc8c6501dddcdea}{
\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} \textbf{\_\-result}}
\label{class_threshold_f5e82e77ffad10da1dc8c6501dddcdea}

\item 
\hypertarget{class_threshold_cf3f8f39e3d71732c76525d62db39116}{
const \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{retain}}
\label{class_threshold_cf3f8f39e3d71732c76525d62db39116}

\item 
\hypertarget{class_threshold_0555939aa91492c2d58b8c650a992e77}{
const \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_threshold_0555939aa91492c2d58b8c650a992e77}

\end{CompactItemize}

\subsection{Detailed Description}
This class takes an Integer array to a boolean array based on a cut-off value.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/thresh.cpp\end{CompactItemize}