### 2009-08-02 doc/latex/files.tex

\section{File List}
Here is a list of all documented files with brief descriptions:\begin{CompactList}
\item\contentsline{section}{cowichan/\hyperlink{cowichan_8cpp}{cowichan.cpp} (Implementation for \hyperlink{class_cowichan}{Cowichan} class and common routines for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan_8cpp}}{}
\item\contentsline{section}{cowichan/\hyperlink{cowichan_8hpp}{cowichan.hpp} (Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan_8hpp}}{}
\item\contentsline{section}{cowichan/\hyperlink{cowichan__defaults_8hpp}{cowichan\_\-defaults.hpp} (Defaults for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan__defaults_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_8cpp}{cowichan\_\-openmp.cpp} (This file contains main method that drives the OpenMP implementation )}{\pageref{cowichan__openmp_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_8hpp}{cowichan\_\-openmp.hpp} (Open Multi-Processing (OpenMP) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__openmp_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2gauss_8cpp}{gauss.cpp} (OpenMP gauss implementation )}{\pageref{cowichan__openmp_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2half_8cpp}{half.cpp} (OpenMP half implementation )}{\pageref{cowichan__openmp_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2hull_8cpp}{hull.cpp} (OpenMP hull implementation )}{\pageref{cowichan__openmp_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2invperc_8cpp}{invperc.cpp} (OpenMP (NOT parallel) invperc implementation )}{\pageref{cowichan__openmp_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2life_8cpp}{life.cpp} (OpenMP life implementation )}{\pageref{cowichan__openmp_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2mandel_8cpp}{mandel.cpp} (OpenMP mandel implementation )}{\pageref{cowichan__openmp_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2norm_8cpp}{norm.cpp} (OpenMP norm implementation )}{\pageref{cowichan__openmp_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2outer_8cpp}{outer.cpp} (OpenMP outer implementation )}{\pageref{cowichan__openmp_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2product_8cpp}{product.cpp} (OpenMP product implementation )}{\pageref{cowichan__openmp_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2randmat_8cpp}{randmat.cpp} (OpenMP randmat implementation )}{\pageref{cowichan__openmp_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sor_8cpp}{sor.cpp} (OpenMP sor implementation )}{\pageref{cowichan__openmp_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{sort_8cpp}{sort.cpp} (Implementation of OpenMP sorting algorithms )}{\pageref{sort_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{sort_8hpp}{sort.hpp} (Sorting algorithms that use OpenMP )}{\pageref{sort_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2thresh_8cpp}{thresh.cpp} (OpenMP thresh implementation )}{\pageref{cowichan__openmp_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2vecdiff_8cpp}{vecdiff.cpp} (OpenMP vecdiff implementation )}{\pageref{cowichan__openmp_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2winnow_8cpp}{winnow.cpp} (OpenMP winnow implementation )}{\pageref{cowichan__openmp_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_8cpp}{cowichan\_\-serial.cpp} (This file contains main method that drives the serial implementation )}{\pageref{cowichan__serial_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_8hpp}{cowichan\_\-serial.hpp} (Serial implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__serial_8hpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2gauss_8cpp}{gauss.cpp} (Serial gauss implementation )}{\pageref{cowichan__serial_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2half_8cpp}{half.cpp} (Serial half implementation )}{\pageref{cowichan__serial_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2hull_8cpp}{hull.cpp} (Serial hull implementation )}{\pageref{cowichan__serial_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2invperc_8cpp}{invperc.cpp} (Serial invperc implementation )}{\pageref{cowichan__serial_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2life_8cpp}{life.cpp} (Serial life implementation )}{\pageref{cowichan__serial_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2mandel_8cpp}{mandel.cpp} (Serial mandel implementation )}{\pageref{cowichan__serial_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2norm_8cpp}{norm.cpp} (Serial norm implementation )}{\pageref{cowichan__serial_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2outer_8cpp}{outer.cpp} (Serial outer implementation )}{\pageref{cowichan__serial_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2product_8cpp}{product.cpp} (Serial product implementation )}{\pageref{cowichan__serial_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2randmat_8cpp}{randmat.cpp} (Serial randmat implementation )}{\pageref{cowichan__serial_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2sor_8cpp}{sor.cpp} (Serial sor implementation )}{\pageref{cowichan__serial_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2thresh_8cpp}{thresh.cpp} (Serial thresh implementation )}{\pageref{cowichan__serial_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2vecdiff_8cpp}{vecdiff.cpp} (Serial vecdiff implementation )}{\pageref{cowichan__serial_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2winnow_8cpp}{winnow.cpp} (Serial winnow implementation )}{\pageref{cowichan__serial_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_8cpp}{cowichan\_\-tbb.cpp} (This file contains main method that drives the TBB implementation )}{\pageref{cowichan__tbb_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_8hpp}{cowichan\_\-tbb.hpp} (Thread Building Blocks (TBB) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__tbb_8hpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2gauss_8cpp}{gauss.cpp} (TBB gauss implementation )}{\pageref{cowichan__tbb_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2half_8cpp}{half.cpp} (TBB half implementation )}{\pageref{cowichan__tbb_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2hull_8cpp}{hull.cpp} (TBB hull implementation )}{\pageref{cowichan__tbb_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2invperc_8cpp}{invperc.cpp} (TBB (NOT parallel) invperc implementation )}{\pageref{cowichan__tbb_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2life_8cpp}{life.cpp} (TBB life implementation )}{\pageref{cowichan__tbb_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2mandel_8cpp}{mandel.cpp} (TBB mandel implementation )}{\pageref{cowichan__tbb_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2norm_8cpp}{norm.cpp} (TBB norm implementation )}{\pageref{cowichan__tbb_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2outer_8cpp}{outer.cpp} (TBB outer implementation )}{\pageref{cowichan__tbb_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2product_8cpp}{product.cpp} (TBB product implementation )}{\pageref{cowichan__tbb_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2randmat_8cpp}{randmat.cpp} (TBB randmat implementation )}{\pageref{cowichan__tbb_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2sor_8cpp}{sor.cpp} (TBB sor implementation )}{\pageref{cowichan__tbb_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2thresh_8cpp}{thresh.cpp} (TBB thresh implementation )}{\pageref{cowichan__tbb_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2vecdiff_8cpp}{vecdiff.cpp} (TBB vecdiff implementation )}{\pageref{cowichan__tbb_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2winnow_8cpp}{winnow.cpp} (TBB winnow implementation )}{\pageref{cowichan__tbb_2winnow_8cpp}}{}
\end{CompactList}