### 2009-08-02 doc/latex/namespacecowichan__openmp.tex

\hypertarget{namespacecowichan__openmp}{
\section{cowichan\_\-openmp Namespace Reference}
\label{namespacecowichan__openmp}\index{cowichan\_\-openmp@{cowichan\_\-openmp}}
}
Additional classes and functions specific to openmp implementation.

\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__openmp_1_1_perc_point}{PercPoint}
\begin{CompactList}\small\item\em Percolation point. \item\end{CompactList}\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_f261011870b201d556308f43b5b25b89}{quickhull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn)
\item 
void \hyperlink{namespacecowichan__openmp_d31787f1b4c9d57119160411e96f769f}{split} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn, \hyperlink{class_point}{Point} $\ast$p1, \hyperlink{class_point}{Point} $\ast$p2)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__openmp_a4ce5b0ba78cd9480bf45e7c313f59d1}{sumNeighbours} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} first, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} r, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} c, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}{mandel\_\-calc} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} x, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} y)
\item 
void \hyperlink{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}{findMinMax} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{Point} $\ast$minPoint, \hyperlink{class_point}{Point} $\ast$maxPoint)
\item 
void \hyperlink{namespacecowichan__openmp_9cbb3540b84f4e051f072a08c2c1d5db}{histogram\_\-sort} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len)
\item 
void \hyperlink{namespacecowichan__openmp_64208258cab6e46db879b46b40581134}{quick\_\-sort} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__openmp_2733c307de545e237e020a1f166850c5}{quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\item 
void \hyperlink{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}{mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$buckets)
\end{CompactItemize}

\subsection{Detailed Description}
Additional classes and functions specific to openmp implementation.

\subsection{Function Documentation}
\hypertarget{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!findMinMax@{findMinMax}}
\index{findMinMax@{findMinMax}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{findMinMax}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::findMinMax ({\bf PointVector} {\em points}, \/  {\bf index\_\-t} {\em n}, \/  {\bf Point} $\ast$ {\em minPoint}, \/  {\bf Point} $\ast$ {\em maxPoint})}}
\label{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}

Find min/max x/y coordinates. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]point vector. \item[{\em n}]number of points. \item[{\em minPoint}]min x/y values. \item[{\em maxPoint}]max x/y values. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_9cbb3540b84f4e051f072a08c2c1d5db}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!histogram\_\-sort@{histogram\_\-sort}}
\index{histogram\_\-sort@{histogram\_\-sort}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{histogram\_\-sort}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::histogram\_\-sort ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len})}}
\label{namespacecowichan__openmp_9cbb3540b84f4e051f072a08c2c1d5db}

This performs parallel histogram sort. \hypertarget{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!mandel\_\-calc@{mandel\_\-calc}}
\index{mandel\_\-calc@{mandel\_\-calc}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{mandel\_\-calc}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} cowichan\_\-openmp::mandel\_\-calc ({\bf real} {\em x}, \/  {\bf real} {\em y})}}
\label{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}

Calculate mandelbrot value. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em x}]x-coordinate. \item[{\em y}]y-coordinate. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Mandelbrot value. \end{Desc}
\hypertarget{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!mask\_\-count@{mask\_\-count}}
\index{mask\_\-count@{mask\_\-count}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{mask\_\-count}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::mask\_\-count ({\bf BoolMatrix} {\em mask}, \/  {\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc}, \/  {\bf index\_\-t} $\ast$ {\em buckets})}}
\label{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}

Count the number of set cells in each thread bucket. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em mask}]boolean mask. \item[{\em nr}]number of rows in mask. \item[{\em nc}]number of columns in mask. \item[{\em buckets}]number of cells in each bucket. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_64208258cab6e46db879b46b40581134}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!quick\_\-sort@{quick\_\-sort}}
\index{quick\_\-sort@{quick\_\-sort}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{quick\_\-sort}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::quick\_\-sort ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len})}}
\label{namespacecowichan__openmp_64208258cab6e46db879b46b40581134}

This performs parallel quick sort. \hypertarget{namespacecowichan__openmp_2733c307de545e237e020a1f166850c5}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}}
\index{quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{quick\_\-sort\_\-partition}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-openmp::quick\_\-sort\_\-partition ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len}, \/  {\bf index\_\-t} {\em pivotIndex})}}
\label{namespacecowichan__openmp_2733c307de545e237e020a1f166850c5}

Partitions vector into points less than and greater than the pivot. \hypertarget{namespacecowichan__openmp_f261011870b201d556308f43b5b25b89}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!quickhull@{quickhull}}
\index{quickhull@{quickhull}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{quickhull}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::quickhull ({\bf PointVector} {\em pointsIn}, \/  {\bf index\_\-t} {\em n}, \/  {\bf PointVector} {\em pointsOut}, \/  {\bf index\_\-t} $\ast$ {\em hn})}}
\label{namespacecowichan__openmp_f261011870b201d556308f43b5b25b89}

Runs quickhull algorithm. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]input points. \item[{\em n}]number of inputs points to use. \item[{\em pointsOut}]output points. \item[{\em hn}]number of output points generated so far. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_d31787f1b4c9d57119160411e96f769f}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!split@{split}}
\index{split@{split}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{split}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::split ({\bf PointVector} {\em pointsIn}, \/  {\bf index\_\-t} {\em n}, \/  {\bf PointVector} {\em pointsOut}, \/  {\bf index\_\-t} $\ast$ {\em hn}, \/  {\bf Point} $\ast$ {\em p1}, \/  {\bf Point} $\ast$ {\em p2})}}
\label{namespacecowichan__openmp_d31787f1b4c9d57119160411e96f769f}

Recursive step of the quickhull algorithm - compute hull on one side of the splitting line. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]input points. \item[{\em n}]number of inputs points to use. \item[{\em pointsOut}]output points. \item[{\em hn}]number of output points generated so far. \item[{\em p1}]first point of the splitting line (p1,p2). \item[{\em p2}]second point of the splitting line (p1,p2). \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_a4ce5b0ba78cd9480bf45e7c313f59d1}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!sumNeighbours@{sumNeighbours}}
\index{sumNeighbours@{sumNeighbours}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{sumNeighbours}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-openmp::sumNeighbours ({\bf BoolMatrix} {\em first}, \/  {\bf index\_\-t} {\em r}, \/  {\bf index\_\-t} {\em c}, \/  {\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc})}}
\label{namespacecowichan__openmp_a4ce5b0ba78cd9480bf45e7c313f59d1}

Calculate number of peers. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em first}]world matrix. \item[{\em r}]row. \item[{\em c}]column. \item[{\em nr}]number of rows in the matrix. \item[{\em nc}]number of columns in the matrix. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]The number of peers. \end{Desc}