### 2009-08-02 doc/latex/namespaces.tex

\section{Namespace List}
Here is a list of all documented namespaces with brief descriptions:\begin{CompactList}
\item\contentsline{section}{\hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp} (Additional classes and functions specific to openmp implementation )}{\pageref{namespacecowichan__openmp}}{}
\item\contentsline{section}{\hyperlink{namespacecowichan__serial}{cowichan\_\-serial} (Additional classes and functions specific to serial implementation )}{\pageref{namespacecowichan__serial}}{}
\item\contentsline{section}{\hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb} (Additional classes and functions specific to tbb implementation )}{\pageref{namespacecowichan__tbb}}{}
\end{CompactList}