### 2009-08-05 doc/latex/namespacecowichan__mpi.tex

\hypertarget{namespacecowichan__mpi}{
\section{cowichan\_\-mpi Namespace Reference}
\label{namespacecowichan__mpi}\index{cowichan\_\-mpi@{cowichan\_\-mpi}}
}
Additional classes and functions specific to mpi implementation.

\subsection*{Functions}
\begin{CompactItemize}
\item 
bool \hyperlink{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}{get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end)
\item 
bool \hyperlink{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}{get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} rank)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}{getNumRemainingNodes} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}{power2} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)
\item 
void \hyperlink{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}{quick\_\-sort} (const mpi::communicator \&world, \hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} owner)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}{quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a54004a105356573d4bac92e1b234aa2}{mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
Additional classes and functions specific to mpi implementation.

\subsection{Function Documentation}
\hypertarget{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!get\_\-block@{get\_\-block}}
\index{get\_\-block@{get\_\-block}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{get\_\-block}]{\setlength{\rightskip}{0pt plus 5cm}bool cowichan\_\-mpi::get\_\-block (const mpi::communicator \& {\em world}, \/  {\bf index\_\-t} {\em lo}, \/  {\bf index\_\-t} {\em hi}, \/  {\bf index\_\-t} $\ast$ {\em start}, \/  {\bf index\_\-t} $\ast$ {\em end}, \/  {\bf index\_\-t} {\em rank})}}
\label{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}

Get a block (start, end) to work on in the range (lo, hi) for process rank. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]communicator. \item[{\em lo}]lower end of the range. \item[{\em hi}]higher end of the range. \item[{\em start}]first index to work on. \item[{\em end}]last index to work on. \item[{\em rank}]process to get work for. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Whether any work is available for process rank. \end{Desc}
\hypertarget{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!get\_\-block@{get\_\-block}}
\index{get\_\-block@{get\_\-block}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{get\_\-block}]{\setlength{\rightskip}{0pt plus 5cm}bool cowichan\_\-mpi::get\_\-block (const mpi::communicator \& {\em world}, \/  {\bf index\_\-t} {\em lo}, \/  {\bf index\_\-t} {\em hi}, \/  {\bf index\_\-t} $\ast$ {\em start}, \/  {\bf index\_\-t} $\ast$ {\em end})}}
\label{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}

Get a block (start, end) to work on in the range (lo, hi) for the current process. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]communicator. \item[{\em lo}]lower end of the range. \item[{\em hi}]higher end of the range. \item[{\em start}]first index to work on. \item[{\em end}]last index to work on. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Whether any work is available for the current process. \end{Desc}
\hypertarget{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!getNumRemainingNodes@{getNumRemainingNodes}}
\index{getNumRemainingNodes@{getNumRemainingNodes}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{getNumRemainingNodes}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::getNumRemainingNodes (const mpi::communicator \& {\em world}, \/  {\bf index\_\-t} {\em level})}}
\label{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}

Get number of remaining nodes, i.e. not doing useful work. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]global communicator. \item[{\em level}]current recursion depth. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Number of remaining nodes. \end{Desc}
\hypertarget{namespacecowichan__mpi_a54004a105356573d4bac92e1b234aa2}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!mask\_\-count@{mask\_\-count}}
\index{mask\_\-count@{mask\_\-count}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{mask\_\-count}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::mask\_\-count ({\bf BoolMatrix} {\em mask}, \/  {\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc})}}
\label{namespacecowichan__mpi_a54004a105356573d4bac92e1b234aa2}

Count the number of set cells in the mask. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em mask}]boolean mask. \item[{\em nr}]number of rows. \item[{\em nc}]number of columns. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Number of set cells in the mask. \end{Desc}
\hypertarget{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!power2@{power2}}
\index{power2@{power2}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{power2}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::power2 ({\bf index\_\-t} {\em n})}}
\label{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}

Calculates a power of 2. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em n}]power is limited by sizeof(index\_\-t) $\ast$ 8 - 2 to prevent overflow. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]2$^\wedge$n. \end{Desc}
\hypertarget{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!quick\_\-sort@{quick\_\-sort}}
\index{quick\_\-sort@{quick\_\-sort}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{quick\_\-sort}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-mpi::quick\_\-sort (const mpi::communicator \& {\em world}, \/  {\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len}, \/  {\bf index\_\-t} {\em level} = {\tt 0}, \/  {\bf index\_\-t} {\em owner} = {\tt 0})}}
\label{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}

This performs parallel quick sort. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]global communicator. \item[{\em vector}]vector to sort. \item[{\em len}]vector length. \item[{\em level}]depth of recursion. \item[{\em owner}]rank of the owner of this sort. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}}
\index{quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{quick\_\-sort\_\-partition}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::quick\_\-sort\_\-partition ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len}, \/  {\bf index\_\-t} {\em pivotIndex})}}
\label{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}

Partitions vector into points less than and greater than the pivot. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em vector}]vector to partition. \item[{\em len}]vector length. \item[{\em pivotIndex}]index to partition around. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]New pivot index. \end{Desc}