### 2009-08-07 doc/latex/cowichan_8hpp.tex

\hypertarget{cowichan_8hpp}{
\section{cowichan/cowichan.hpp File Reference}
\label{cowichan_8hpp}\index{cowichan/cowichan.hpp@{cowichan/cowichan.hpp}}
}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs.

{\tt \#include \char`\"{}cowichan\_\-defaults.hpp\char`\"{}}\par
{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdlib$>$}\par
{\tt \#include $<$ctime$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include $<$algorithm$>$}\par
{\tt \#include $<$iomanip$>$}\par
{\tt \#include $<$climits$>$}\par
{\tt \#include $<$limits$>$}\par
{\tt \#include $<$string$>$}\par
{\tt \#include $<$cstring$>$}\par
{\tt \#include $<$windows.h$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_point}{Point}
\begin{CompactList}\small\item\em Two-dimensional point. \item\end{CompactList}\item 
class \hyperlink{class_weighted_point}{WeightedPoint}
\begin{CompactList}\small\item\em Two-dimensional point with an integer weight. \item\end{CompactList}\item 
class \hyperlink{class_cowichan}{Cowichan}
\begin{CompactList}\small\item\em Base class for all C++ implementations. \item\end{CompactList}\end{CompactItemize}
\subsection*{Defines}
\begin{CompactItemize}
\item 
\#define \hyperlink{cowichan_8hpp_9adcfa1dcb92dd0d8757264ae41d1710}{TEST\_\-TIME}
\item 
\#define \hyperlink{cowichan_8hpp_e0937999ad4f52ae055b86396280604a}{SORT\_\-TIME}
\item 
\#define \hyperlink{cowichan_8hpp_7db5e8ee7c27e6fb5c9caf9d674ea949}{REAL\_\-TYPE}~float
\item 
\#define \hyperlink{cowichan_8hpp_9f8bbf20bc9432260a35b2f3cdb009f8}{MAXIMUM\_\-INT}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE}$>$::max()
\item 
\#define \hyperlink{cowichan_8hpp_af258621cee3ac54ad546ab5297c2476}{MINIMUM\_\-INT}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE}$>$::min()
\item 
\#define \hyperlink{cowichan_8hpp_753345ec3932dd32057b98ab9386efa0}{MAXIMUM\_\-REAL}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}$>$::min()
\item 
\#define \hyperlink{cowichan_8hpp_e618a02a4bb148a144f9b9f99aa0b581}{MINIMUM\_\-REAL}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}$>$::min()
\item 
\#define \hyperlink{cowichan_8hpp_bb50b5f693a0b64cdf91bca21cb464de}{INFINITY\_\-REAL}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}$>$::infinity()
\item 
\#define \hyperlink{cowichan_8hpp_e2f68e53d100109588bb54ccc1c3d85b}{MATRIX\_\-RECT}(mtrx, row, col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ nc + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_2747f4d7d371c6ebdcaeb134876a2cf3}{MATRIX\_\-RECT\_\-NC}(mtrx, row, col, nc)~(mtrx)\mbox{[}(row)$\ast$(nc) + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_79012ec47cbea66cd6664a54ca8a754c}{MATRIX\_\-SQUARE}(mtrx, row, col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ n + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_6772060ae1a2d5750934131acf95322a}{MATRIX\_\-SQUARE\_\-N}(mtrx, row, col, n)~(mtrx)\mbox{[}(row)$\ast$(n) + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_f1d98fb728b5c7300f80dd782702d1dd}{MATRIX}~MATRIX\_\-SQUARE
\item 
\#define \hyperlink{cowichan_8hpp_41a3c329b378fa03397ed8a8925cbc73}{VECTOR}(vect, row)~(vect)\mbox{[}row\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_bacaf90ccfec9927a6d97aa37f39917d}{DIAG}(mtrx, v)~(mtrx)\mbox{[}(v)$\ast$this $\rightarrow$ n + v\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_7b818ef6ccf7d45bd25b07a402ca500f}{DIAG\_\-SZ}(mtrx, v, n)~(mtrx)\mbox{[}(v)$\ast$(n) + v\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_07e1668b7974a041604d27b62cbe76f5}{NEW\_\-MATRIX\_\-RECT}(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ nr $\ast$ this $\rightarrow$ nc\mbox{]})
\item 
\#define \hyperlink{cowichan_8hpp_03c9abf69e2e60f4ed74eebad85b750f}{NEW\_\-MATRIX\_\-SQUARE}(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ n $\ast$ this $\rightarrow$ n\mbox{]})
\item 
\#define \hyperlink{cowichan_8hpp_e37a7bc1a3de6387853e097f2a8b24b5}{NEW\_\-VECTOR\_\-SZ}(\_\-\_\-type, \_\-\_\-num)~(new \_\-\_\-type\mbox{[}\_\-\_\-num\mbox{]})
\item 
\#define \hyperlink{cowichan_8hpp_71e0e9927a2730d38d0335484f7554fe}{NEW\_\-VECTOR}(\_\-\_\-type)~NEW\_\-VECTOR\_\-SZ(\_\-\_\-type, this $\rightarrow$ n)
\end{CompactItemize}
\subsection*{Typedefs}
\begin{CompactItemize}
\item 
typedef REAL\_\-TYPE \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}
\item 
typedef UINT32 \hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE}
\item 
typedef \hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} $\ast$ \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix}
\item 
typedef bool $\ast$ \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix}
\item 
typedef \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} $\ast$ \hyperlink{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}{RealMatrix}
\item 
typedef \hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} $\ast$ \hyperlink{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}{IntVector}
\item 
typedef bool $\ast$ \hyperlink{cowichan_8hpp_69facfadf10e8e6e1b3ea0d02af7fef4}{BoolVector}
\item 
typedef \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} $\ast$ \hyperlink{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}{RealVector}
\item 
typedef \hyperlink{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}{RealMatrix} \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix}
\item 
typedef \hyperlink{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}{RealVector} \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector}
\item 
typedef ptrdiff\_\-t \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t}
\item 
typedef \hyperlink{class_point}{Point} $\ast$ \hyperlink{cowichan_8hpp_1551f7def82edc60d7e37053954b7a07}{PointVector}
\item 
typedef \hyperlink{class_weighted_point}{WeightedPoint} $\ast$ \hyperlink{cowichan_8hpp_bd6eb1c54f195e0982817565e2ca9c31}{WeightedPointVector}
\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
INT64 \hyperlink{cowichan_8hpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}{get\_\-ticks} ()
\item 
INT64 \hyperlink{cowichan_8hpp_40356e6c8804a8f94e41c156ad368c83}{get\_\-freq} ()
\item 
void \hyperlink{cowichan_8hpp_560007e135455d06968b8a834d970cf9}{print\_\-elapsed\_\-time} (INT64 start, INT64 end)
\item 
void \hyperlink{cowichan_8hpp_1bf3d816eb607b5a06a97e05e9b95099}{timeInfo} (INT64 $\ast$start, INT64 $\ast$end, std::string message)
\item 
void \hyperlink{cowichan_8hpp_faa5a0a08b60255e5a59cbd3b1f5e45f}{out\_\-of\_\-memory} ()
\item 
void \hyperlink{cowichan_8hpp_501801b9dcd90a910432c7712ff763e7}{not\_\-enough\_\-points} ()
\item 
void \hyperlink{cowichan_8hpp_19e1551f7e954a984729d96c2c0bf5ba}{no\_\-cells\_\-alive} ()
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{cowichan_8hpp_eb2386a0fab0a548605cd5f0898da5b4}{uniform} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} mean, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} range)
\end{CompactItemize}

\subsection{Detailed Description}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs.

\subsection{Define Documentation}
\hypertarget{cowichan_8hpp_bacaf90ccfec9927a6d97aa37f39917d}{
\index{cowichan.hpp@{cowichan.hpp}!DIAG@{DIAG}}
\index{DIAG@{DIAG}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{DIAG}]{\setlength{\rightskip}{0pt plus 5cm}\#define DIAG(mtrx, \/  v)~(mtrx)\mbox{[}(v)$\ast$this $\rightarrow$ n + v\mbox{]}}}
\label{cowichan_8hpp_bacaf90ccfec9927a6d97aa37f39917d}

Access a square matrix diagonal element v from a class using member variable n as the matrix size. \hypertarget{cowichan_8hpp_7b818ef6ccf7d45bd25b07a402ca500f}{
\index{cowichan.hpp@{cowichan.hpp}!DIAG\_\-SZ@{DIAG\_\-SZ}}
\index{DIAG\_\-SZ@{DIAG\_\-SZ}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{DIAG\_\-SZ}]{\setlength{\rightskip}{0pt plus 5cm}\#define DIAG\_\-SZ(mtrx, \/  v, \/  n)~(mtrx)\mbox{[}(v)$\ast$(n) + v\mbox{]}}}
\label{cowichan_8hpp_7b818ef6ccf7d45bd25b07a402ca500f}

Access a square matrix diagonal element v using n as the matrix size. \hypertarget{cowichan_8hpp_bb50b5f693a0b64cdf91bca21cb464de}{
\index{cowichan.hpp@{cowichan.hpp}!INFINITY\_\-REAL@{INFINITY\_\-REAL}}
\index{INFINITY\_\-REAL@{INFINITY\_\-REAL}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{INFINITY\_\-REAL}]{\setlength{\rightskip}{0pt plus 5cm}\#define INFINITY\_\-REAL~numeric\_\-limits$<${\bf real}$>$::infinity()}}
\label{cowichan_8hpp_bb50b5f693a0b64cdf91bca21cb464de}

Real infinity value. \hypertarget{cowichan_8hpp_f1d98fb728b5c7300f80dd782702d1dd}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX@{MATRIX}}
\index{MATRIX@{MATRIX}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX~MATRIX\_\-SQUARE}}
\label{cowichan_8hpp_f1d98fb728b5c7300f80dd782702d1dd}

Same as MATRIX\_\-SQUARE (shorter name) \hypertarget{cowichan_8hpp_e2f68e53d100109588bb54ccc1c3d85b}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-RECT@{MATRIX\_\-RECT}}
\index{MATRIX\_\-RECT@{MATRIX\_\-RECT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-RECT}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-RECT(mtrx, \/  row, \/  col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ nc + col\mbox{]}}}
\label{cowichan_8hpp_e2f68e53d100109588bb54ccc1c3d85b}

Access a rectangular matrix at row/col from a class using member variable nc as the number of columns. \hypertarget{cowichan_8hpp_2747f4d7d371c6ebdcaeb134876a2cf3}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-RECT\_\-NC@{MATRIX\_\-RECT\_\-NC}}
\index{MATRIX\_\-RECT\_\-NC@{MATRIX\_\-RECT\_\-NC}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-RECT\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-RECT\_\-NC(mtrx, \/  row, \/  col, \/  nc)~(mtrx)\mbox{[}(row)$\ast$(nc) + col\mbox{]}}}
\label{cowichan_8hpp_2747f4d7d371c6ebdcaeb134876a2cf3}

Access a rectangular matrix at row/col using nc as the number of columns. \hypertarget{cowichan_8hpp_79012ec47cbea66cd6664a54ca8a754c}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-SQUARE@{MATRIX\_\-SQUARE}}
\index{MATRIX\_\-SQUARE@{MATRIX\_\-SQUARE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-SQUARE}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-SQUARE(mtrx, \/  row, \/  col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ n + col\mbox{]}}}
\label{cowichan_8hpp_79012ec47cbea66cd6664a54ca8a754c}

Access a square matrix at row/col from a class using member variable n as the matrix size. \hypertarget{cowichan_8hpp_6772060ae1a2d5750934131acf95322a}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-SQUARE\_\-N@{MATRIX\_\-SQUARE\_\-N}}
\index{MATRIX\_\-SQUARE\_\-N@{MATRIX\_\-SQUARE\_\-N}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-SQUARE\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-SQUARE\_\-N(mtrx, \/  row, \/  col, \/  n)~(mtrx)\mbox{[}(row)$\ast$(n) + col\mbox{]}}}
\label{cowichan_8hpp_6772060ae1a2d5750934131acf95322a}

Access a square matrix at row/col using n as the matrix size. \hypertarget{cowichan_8hpp_9f8bbf20bc9432260a35b2f3cdb009f8}{
\index{cowichan.hpp@{cowichan.hpp}!MAXIMUM\_\-INT@{MAXIMUM\_\-INT}}
\index{MAXIMUM\_\-INT@{MAXIMUM\_\-INT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MAXIMUM\_\-INT}]{\setlength{\rightskip}{0pt plus 5cm}\#define MAXIMUM\_\-INT~numeric\_\-limits$<${\bf INT\_\-TYPE}$>$::max()}}
\label{cowichan_8hpp_9f8bbf20bc9432260a35b2f3cdb009f8}

Maximum integer value in matrices/vectors. \hypertarget{cowichan_8hpp_753345ec3932dd32057b98ab9386efa0}{
\index{cowichan.hpp@{cowichan.hpp}!MAXIMUM\_\-REAL@{MAXIMUM\_\-REAL}}
\index{MAXIMUM\_\-REAL@{MAXIMUM\_\-REAL}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MAXIMUM\_\-REAL}]{\setlength{\rightskip}{0pt plus 5cm}\#define MAXIMUM\_\-REAL~numeric\_\-limits$<${\bf real}$>$::min()}}
\label{cowichan_8hpp_753345ec3932dd32057b98ab9386efa0}

Maximum real value in matrices/vectors. \hypertarget{cowichan_8hpp_af258621cee3ac54ad546ab5297c2476}{
\index{cowichan.hpp@{cowichan.hpp}!MINIMUM\_\-INT@{MINIMUM\_\-INT}}
\index{MINIMUM\_\-INT@{MINIMUM\_\-INT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MINIMUM\_\-INT}]{\setlength{\rightskip}{0pt plus 5cm}\#define MINIMUM\_\-INT~numeric\_\-limits$<${\bf INT\_\-TYPE}$>$::min()}}
\label{cowichan_8hpp_af258621cee3ac54ad546ab5297c2476}

Minimum integer value in matrices/vectors. \hypertarget{cowichan_8hpp_e618a02a4bb148a144f9b9f99aa0b581}{
\index{cowichan.hpp@{cowichan.hpp}!MINIMUM\_\-REAL@{MINIMUM\_\-REAL}}
\index{MINIMUM\_\-REAL@{MINIMUM\_\-REAL}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MINIMUM\_\-REAL}]{\setlength{\rightskip}{0pt plus 5cm}\#define MINIMUM\_\-REAL~numeric\_\-limits$<${\bf real}$>$::min()}}
\label{cowichan_8hpp_e618a02a4bb148a144f9b9f99aa0b581}

Minimum real value in matrices/vectors. \hypertarget{cowichan_8hpp_07e1668b7974a041604d27b62cbe76f5}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-MATRIX\_\-RECT@{NEW\_\-MATRIX\_\-RECT}}
\index{NEW\_\-MATRIX\_\-RECT@{NEW\_\-MATRIX\_\-RECT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-MATRIX\_\-RECT}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-MATRIX\_\-RECT(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ nr $\ast$ this $\rightarrow$ nc\mbox{]})}}
\label{cowichan_8hpp_07e1668b7974a041604d27b62cbe76f5}

Allocate a new rectangular matrix from a class using member variable nr as the number of rows and member variable nc as the number of columns. \hypertarget{cowichan_8hpp_03c9abf69e2e60f4ed74eebad85b750f}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-MATRIX\_\-SQUARE@{NEW\_\-MATRIX\_\-SQUARE}}
\index{NEW\_\-MATRIX\_\-SQUARE@{NEW\_\-MATRIX\_\-SQUARE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-MATRIX\_\-SQUARE}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-MATRIX\_\-SQUARE(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ n $\ast$ this $\rightarrow$ n\mbox{]})}}
\label{cowichan_8hpp_03c9abf69e2e60f4ed74eebad85b750f}

Allocate a new square matrix from a class using member variable n as the matrix size. \hypertarget{cowichan_8hpp_71e0e9927a2730d38d0335484f7554fe}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-VECTOR@{NEW\_\-VECTOR}}
\index{NEW\_\-VECTOR@{NEW\_\-VECTOR}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-VECTOR}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-VECTOR(\_\-\_\-type)~NEW\_\-VECTOR\_\-SZ(\_\-\_\-type, this $\rightarrow$ n)}}
\label{cowichan_8hpp_71e0e9927a2730d38d0335484f7554fe}

Allocate a new vector from a class using member variable n as the size of the vector. \hypertarget{cowichan_8hpp_e37a7bc1a3de6387853e097f2a8b24b5}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-VECTOR\_\-SZ@{NEW\_\-VECTOR\_\-SZ}}
\index{NEW\_\-VECTOR\_\-SZ@{NEW\_\-VECTOR\_\-SZ}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-VECTOR\_\-SZ}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-VECTOR\_\-SZ(\_\-\_\-type, \/  \_\-\_\-num)~(new \_\-\_\-type\mbox{[}\_\-\_\-num\mbox{]})}}
\label{cowichan_8hpp_e37a7bc1a3de6387853e097f2a8b24b5}

Allocate a new vector of size \_\-\_\-num. \hypertarget{cowichan_8hpp_7db5e8ee7c27e6fb5c9caf9d674ea949}{
\index{cowichan.hpp@{cowichan.hpp}!REAL\_\-TYPE@{REAL\_\-TYPE}}
\index{REAL\_\-TYPE@{REAL\_\-TYPE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{REAL\_\-TYPE}]{\setlength{\rightskip}{0pt plus 5cm}\#define REAL\_\-TYPE~float}}
\label{cowichan_8hpp_7db5e8ee7c27e6fb5c9caf9d674ea949}

Use IEEE single floating-point by default. \hypertarget{cowichan_8hpp_e0937999ad4f52ae055b86396280604a}{
\index{cowichan.hpp@{cowichan.hpp}!SORT\_\-TIME@{SORT\_\-TIME}}
\index{SORT\_\-TIME@{SORT\_\-TIME}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{SORT\_\-TIME}]{\setlength{\rightskip}{0pt plus 5cm}\#define SORT\_\-TIME}}
\label{cowichan_8hpp_e0937999ad4f52ae055b86396280604a}

Enables printing of sort time (in winnow). \hypertarget{cowichan_8hpp_9adcfa1dcb92dd0d8757264ae41d1710}{
\index{cowichan.hpp@{cowichan.hpp}!TEST\_\-TIME@{TEST\_\-TIME}}
\index{TEST\_\-TIME@{TEST\_\-TIME}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{TEST\_\-TIME}]{\setlength{\rightskip}{0pt plus 5cm}\#define TEST\_\-TIME}}
\label{cowichan_8hpp_9adcfa1dcb92dd0d8757264ae41d1710}

Enables printing of test time. \hypertarget{cowichan_8hpp_41a3c329b378fa03397ed8a8925cbc73}{
\index{cowichan.hpp@{cowichan.hpp}!VECTOR@{VECTOR}}
\index{VECTOR@{VECTOR}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{VECTOR}]{\setlength{\rightskip}{0pt plus 5cm}\#define VECTOR(vect, \/  row)~(vect)\mbox{[}row\mbox{]}}}
\label{cowichan_8hpp_41a3c329b378fa03397ed8a8925cbc73}

Access row of a vector.

\subsection{Typedef Documentation}
\hypertarget{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{
\index{cowichan.hpp@{cowichan.hpp}!BoolMatrix@{BoolMatrix}}
\index{BoolMatrix@{BoolMatrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{BoolMatrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef bool$\ast$ {\bf BoolMatrix}}}
\label{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}

Boolean matrix type. \hypertarget{cowichan_8hpp_69facfadf10e8e6e1b3ea0d02af7fef4}{
\index{cowichan.hpp@{cowichan.hpp}!BoolVector@{BoolVector}}
\index{BoolVector@{BoolVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{BoolVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef bool$\ast$ {\bf BoolVector}}}
\label{cowichan_8hpp_69facfadf10e8e6e1b3ea0d02af7fef4}

Boolean vector type. \hypertarget{cowichan_8hpp_5b04577d5d21124855deaad298595371}{
\index{cowichan.hpp@{cowichan.hpp}!index\_\-t@{index\_\-t}}
\index{index\_\-t@{index\_\-t}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{index\_\-t}]{\setlength{\rightskip}{0pt plus 5cm}typedef ptrdiff\_\-t {\bf index\_\-t}}}
\label{cowichan_8hpp_5b04577d5d21124855deaad298595371}

Index type (signed version of size\_\-t). \hypertarget{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{
\index{cowichan.hpp@{cowichan.hpp}!INT\_\-TYPE@{INT\_\-TYPE}}
\index{INT\_\-TYPE@{INT\_\-TYPE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{INT\_\-TYPE}]{\setlength{\rightskip}{0pt plus 5cm}typedef UINT32 {\bf INT\_\-TYPE}}}
\label{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}

Integer type (used in matrix/vector values). \hypertarget{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{
\index{cowichan.hpp@{cowichan.hpp}!IntMatrix@{IntMatrix}}
\index{IntMatrix@{IntMatrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{IntMatrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf INT\_\-TYPE}$\ast$ {\bf IntMatrix}}}
\label{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}

Integer matrix type. \hypertarget{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}{
\index{cowichan.hpp@{cowichan.hpp}!IntVector@{IntVector}}
\index{IntVector@{IntVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{IntVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf INT\_\-TYPE}$\ast$ {\bf IntVector}}}
\label{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}

Integer vector type. \hypertarget{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{
\index{cowichan.hpp@{cowichan.hpp}!Matrix@{Matrix}}
\index{Matrix@{Matrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{Matrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf RealMatrix} {\bf Matrix}}}
\label{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}

Real matrix type (shorter name). \hypertarget{cowichan_8hpp_1551f7def82edc60d7e37053954b7a07}{
\index{cowichan.hpp@{cowichan.hpp}!PointVector@{PointVector}}
\index{PointVector@{PointVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{PointVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf Point}$\ast$ {\bf PointVector}}}
\label{cowichan_8hpp_1551f7def82edc60d7e37053954b7a07}

Vector of points type. \hypertarget{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{
\index{cowichan.hpp@{cowichan.hpp}!real@{real}}
\index{real@{real}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{real}]{\setlength{\rightskip}{0pt plus 5cm}typedef REAL\_\-TYPE {\bf real}}}
\label{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}

Real type (used in matrix/vector values). \hypertarget{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}{
\index{cowichan.hpp@{cowichan.hpp}!RealMatrix@{RealMatrix}}
\index{RealMatrix@{RealMatrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{RealMatrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf real}$\ast$ {\bf RealMatrix}}}
\label{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}

Real matrix type. \hypertarget{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}{
\index{cowichan.hpp@{cowichan.hpp}!RealVector@{RealVector}}
\index{RealVector@{RealVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{RealVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf real}$\ast$ {\bf RealVector}}}
\label{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}

Real vector type. \hypertarget{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{
\index{cowichan.hpp@{cowichan.hpp}!Vector@{Vector}}
\index{Vector@{Vector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{Vector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf RealVector} {\bf Vector}}}
\label{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}

Real vector type (shorter name). \hypertarget{cowichan_8hpp_bd6eb1c54f195e0982817565e2ca9c31}{
\index{cowichan.hpp@{cowichan.hpp}!WeightedPointVector@{WeightedPointVector}}
\index{WeightedPointVector@{WeightedPointVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{WeightedPointVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf WeightedPoint}$\ast$ {\bf WeightedPointVector}}}
\label{cowichan_8hpp_bd6eb1c54f195e0982817565e2ca9c31}

Vector of weighted points type.

\subsection{Function Documentation}
\hypertarget{cowichan_8hpp_40356e6c8804a8f94e41c156ad368c83}{
\index{cowichan.hpp@{cowichan.hpp}!get\_\-freq@{get\_\-freq}}
\index{get\_\-freq@{get\_\-freq}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{get\_\-freq}]{\setlength{\rightskip}{0pt plus 5cm}INT64 get\_\-freq ()}}
\label{cowichan_8hpp_40356e6c8804a8f94e41c156ad368c83}

Get tick frequency. \begin{Desc}
\item[Returns:]Tick frequency. \end{Desc}
\hypertarget{cowichan_8hpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}{
\index{cowichan.hpp@{cowichan.hpp}!get\_\-ticks@{get\_\-ticks}}
\index{get\_\-ticks@{get\_\-ticks}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{get\_\-ticks}]{\setlength{\rightskip}{0pt plus 5cm}INT64 get\_\-ticks ()}}
\label{cowichan_8hpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}

Get tick count. \begin{Desc}
\item[Returns:]Tick count. \end{Desc}
\hypertarget{cowichan_8hpp_19e1551f7e954a984729d96c2c0bf5ba}{
\index{cowichan.hpp@{cowichan.hpp}!no\_\-cells\_\-alive@{no\_\-cells\_\-alive}}
\index{no\_\-cells\_\-alive@{no\_\-cells\_\-alive}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{no\_\-cells\_\-alive}]{\setlength{\rightskip}{0pt plus 5cm}void no\_\-cells\_\-alive ()}}
\label{cowichan_8hpp_19e1551f7e954a984729d96c2c0bf5ba}

Prints no cells alive message and exits. \hypertarget{cowichan_8hpp_501801b9dcd90a910432c7712ff763e7}{
\index{cowichan.hpp@{cowichan.hpp}!not\_\-enough\_\-points@{not\_\-enough\_\-points}}
\index{not\_\-enough\_\-points@{not\_\-enough\_\-points}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{not\_\-enough\_\-points}]{\setlength{\rightskip}{0pt plus 5cm}void not\_\-enough\_\-points ()}}
\label{cowichan_8hpp_501801b9dcd90a910432c7712ff763e7}

Prints not enough points message and exits. \hypertarget{cowichan_8hpp_faa5a0a08b60255e5a59cbd3b1f5e45f}{
\index{cowichan.hpp@{cowichan.hpp}!out\_\-of\_\-memory@{out\_\-of\_\-memory}}
\index{out\_\-of\_\-memory@{out\_\-of\_\-memory}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{out\_\-of\_\-memory}]{\setlength{\rightskip}{0pt plus 5cm}void out\_\-of\_\-memory ()}}
\label{cowichan_8hpp_faa5a0a08b60255e5a59cbd3b1f5e45f}

Prints out of memory message and exits. \hypertarget{cowichan_8hpp_560007e135455d06968b8a834d970cf9}{
\index{cowichan.hpp@{cowichan.hpp}!print\_\-elapsed\_\-time@{print\_\-elapsed\_\-time}}
\index{print\_\-elapsed\_\-time@{print\_\-elapsed\_\-time}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{print\_\-elapsed\_\-time}]{\setlength{\rightskip}{0pt plus 5cm}void print\_\-elapsed\_\-time (INT64 {\em start}, \/  INT64 {\em end})}}
\label{cowichan_8hpp_560007e135455d06968b8a834d970cf9}

Print elapsed time to std::cout. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em start}]initial tick count. \item[{\em end}]final tick count. \end{description}
\end{Desc}
\hypertarget{cowichan_8hpp_1bf3d816eb607b5a06a97e05e9b95099}{
\index{cowichan.hpp@{cowichan.hpp}!timeInfo@{timeInfo}}
\index{timeInfo@{timeInfo}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{timeInfo}]{\setlength{\rightskip}{0pt plus 5cm}void timeInfo (INT64 $\ast$ {\em start}, \/  INT64 $\ast$ {\em end}, \/  std::string {\em message})}}
\label{cowichan_8hpp_1bf3d816eb607b5a06a97e05e9b95099}

Does a sort of swap-out, printing progress. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em start}]pointer to initial tick count. \item[{\em end}]pointeger to final tick count. \item[{\em message}]message to print before the elapsed time. \end{description}
\end{Desc}
\hypertarget{cowichan_8hpp_eb2386a0fab0a548605cd5f0898da5b4}{
\index{cowichan.hpp@{cowichan.hpp}!uniform@{uniform}}
\index{uniform@{uniform}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{uniform}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} uniform ({\bf real} {\em mean}, \/  {\bf real} {\em range})}}
\label{cowichan_8hpp_eb2386a0fab0a548605cd5f0898da5b4}

Returns a pseudorandom number $\sim$ U\mbox{[}mean - range, mean + range\mbox{]}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em mean}]mean. \item[{\em range}]range. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]The pseudorandom number. \end{Desc}