### 2009-08-07 doc/latex/cowichan__lt_2src_2gauss_8cpp.tex

\hypertarget{cowichan__lt_2src_2gauss_8cpp}{
\section{cowichan\_\-lt/src/gauss.cpp File Reference}
\label{cowichan__lt_2src_2gauss_8cpp}\index{cowichan\_\-lt/src/gauss.cpp@{cowichan\_\-lt/src/gauss.cpp}}
}
LinuxTuples gauss implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}gauss.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples gauss implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_fd3e659afb38415f56fd508333b7bb1b}{CowichanLinuxTuples::gauss} \end{Desc}