### 2009-08-07 doc/latex/cowichan__lt_2src_2half_8cpp.tex

\hypertarget{cowichan__lt_2src_2half_8cpp}{
\section{cowichan\_\-lt/src/half.cpp File Reference}
\label{cowichan__lt_2src_2half_8cpp}\index{cowichan\_\-lt/src/half.cpp@{cowichan\_\-lt/src/half.cpp}}
}
LinuxTuples half implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include \char`\"{}half.hpp\char`\"{}}\par
\subsection*{Defines}
\begin{CompactItemize}
\item 
\hypertarget{cowichan__lt_2src_2half_8cpp_44133fb2267b03d2193d855394486d0f}{
\#define \textbf{FLIP}(item)~(item \% 2 == 0 ? (item + 1) : (item - 1))}
\label{cowichan__lt_2src_2half_8cpp_44133fb2267b03d2193d855394486d0f}

\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples half implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_f3bd58ad2653f06d242d473e7bffda0b}{CowichanLinuxTuples::half} \end{Desc}