### 2009-08-07 doc/latex/cowichan__lt_2src_2hull_8cpp.tex

\hypertarget{cowichan__lt_2src_2hull_8cpp}{
\section{cowichan\_\-lt/src/hull.cpp File Reference}
\label{cowichan__lt_2src_2hull_8cpp}\index{cowichan\_\-lt/src/hull.cpp@{cowichan\_\-lt/src/hull.cpp}}
}
LinuxTuples hull implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}hull.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples hull implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_501f87594d62af261ff0b1954a60ddc4}{CowichanLinuxTuples::hull} \end{Desc}