### 2009-08-07 doc/latex/cowichan__lt_2src_2randmat_8cpp.tex

\hypertarget{cowichan__lt_2src_2randmat_8cpp}{
\section{cowichan\_\-lt/src/randmat.cpp File Reference}
\label{cowichan__lt_2src_2randmat_8cpp}\index{cowichan\_\-lt/src/randmat.cpp@{cowichan\_\-lt/src/randmat.cpp}}
}
LinuxTuples randmat implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include \char`\"{}randmat.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples randmat implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_ff90c6a27db51366570b70d352de7fab}{CowichanLinuxTuples::randmat} \end{Desc}