### 2009-08-07 doc/latex/cowichan__lt_2src_2sor_8cpp.tex

\hypertarget{cowichan__lt_2src_2sor_8cpp}{
\section{cowichan\_\-lt/src/sor.cpp File Reference}
\label{cowichan__lt_2src_2sor_8cpp}\index{cowichan\_\-lt/src/sor.cpp@{cowichan\_\-lt/src/sor.cpp}}
}
LinuxTuples successive over-relaxation implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}sor.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples successive over-relaxation implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_aa2c469c3f520a5a56c5ec1ac9a5a6c6}{CowichanLinuxTuples::sor} \end{Desc}