### 2009-08-07 doc/latex/cowichan__lt_2src_2thresh_8cpp.tex

\hypertarget{cowichan__lt_2src_2thresh_8cpp}{
\section{cowichan\_\-lt/src/thresh.cpp File Reference}
\label{cowichan__lt_2src_2thresh_8cpp}\index{cowichan\_\-lt/src/thresh.cpp@{cowichan\_\-lt/src/thresh.cpp}}
}
LinuxTuples thresh implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$map$>$}\par
{\tt \#include \char`\"{}thresh.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_b8cf055ec3d42a1ccde6d818c5a43a74}{CowichanLinuxTuples::thresh} \end{Desc}