### 2009-08-07 doc/latex/cowichan__lt_8cpp.tex

\hypertarget{cowichan__lt_8cpp}{
\section{cowichan\_\-lt/src/cowichan\_\-lt.cpp File Reference}
\label{cowichan__lt_8cpp}\index{cowichan\_\-lt/src/cowichan\_\-lt.cpp@{cowichan\_\-lt/src/cowichan\_\-lt.cpp}}
}
Entry point for the cowichan linuxtuples application.

{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__lt_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\end{CompactItemize}

\subsection{Detailed Description}
Entry point for the cowichan linuxtuples application.

\subsection{Function Documentation}
\hypertarget{cowichan__lt_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-lt.cpp@{cowichan\_\-lt.cpp}!main@{main}}
\index{main@{main}!cowichan_lt.cpp@{cowichan\_\-lt.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__lt_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Entry point of the program.