### 2009-08-07 doc/latex/cowichan__lt_8hpp.tex

\hypertarget{cowichan__lt_8hpp}{
\section{cowichan\_\-lt/src/cowichan\_\-lt.hpp File Reference}
\label{cowichan__lt_8hpp}\index{cowichan\_\-lt/src/cowichan\_\-lt.hpp@{cowichan\_\-lt/src/cowichan\_\-lt.hpp}}
}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs; LinuxTuples implementation.

{\tt \#include \char`\"{}../../cowichan/cowichan.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_cowichan_linux_tuples}{CowichanLinuxTuples}
\end{CompactItemize}

\subsection{Detailed Description}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs; LinuxTuples implementation.