### 2009-08-07 doc/latex/cowichan__mpi_2sort_8cpp.tex

\hypertarget{cowichan__mpi_2sort_8cpp}{
\section{cowichan\_\-mpi/sort.cpp File Reference}
\label{cowichan__mpi_2sort_8cpp}\index{cowichan\_\-mpi/sort.cpp@{cowichan\_\-mpi/sort.cpp}}
}
Implementation of MPI sorting algorithms.

{\tt \#include \char`\"{}sort.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__mpi}{cowichan\_\-mpi}
\begin{CompactList}\small\item\em Additional classes and functions specific to mpi implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}{cowichan\_\-mpi::getNumRemainingNodes} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}{cowichan\_\-mpi::power2} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)
\item 
void \hyperlink{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}{cowichan\_\-mpi::quick\_\-sort} (const mpi::communicator \&world, \hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} owner)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}{cowichan\_\-mpi::quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\end{CompactItemize}

\subsection{Detailed Description}
Implementation of MPI sorting algorithms.