### 2009-08-07 doc/latex/cowichan__mpi_2sort_8hpp.tex

\hypertarget{cowichan__mpi_2sort_8hpp}{
\section{cowichan\_\-mpi/sort.hpp File Reference}
\label{cowichan__mpi_2sort_8hpp}\index{cowichan\_\-mpi/sort.hpp@{cowichan\_\-mpi/sort.hpp}}
}
Sorting algorithms that use MPI.

{\tt \#include \char`\"{}cowichan\_\-mpi.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__mpi}{cowichan\_\-mpi}
\begin{CompactList}\small\item\em Additional classes and functions specific to mpi implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Defines}
\begin{CompactItemize}
\item 
\#define \hyperlink{cowichan__mpi_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{QUICK\_\-SORT\_\-CUTOFF}~100
\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}{cowichan\_\-mpi::getNumRemainingNodes} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}{cowichan\_\-mpi::power2} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)
\item 
void \hyperlink{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}{cowichan\_\-mpi::quick\_\-sort} (const mpi::communicator \&world, \hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} owner)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}{cowichan\_\-mpi::quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\end{CompactItemize}

\subsection{Detailed Description}
Sorting algorithms that use MPI.

One algorithms is included: \begin{enumerate}
\item Parallel Quick Sort \end{enumerate}

\subsection{Define Documentation}
\hypertarget{cowichan__mpi_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{
\index{cowichan\_\-mpi/sort.hpp@{cowichan\_\-mpi/sort.hpp}!QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}}
\index{QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}!cowichan_mpi/sort.hpp@{cowichan\_\-mpi/sort.hpp}}
\subsubsection[{QUICK\_\-SORT\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-CUTOFF~100}}
\label{cowichan__mpi_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}

Quick sort cutoff (becomes insertion sort) - must be $>$ 1