### 2009-08-07 doc/latex/cowichan__mpi_8cpp.tex

\hypertarget{cowichan__mpi_8cpp}{
\section{cowichan\_\-mpi/cowichan\_\-mpi.cpp File Reference}
\label{cowichan__mpi_8cpp}\index{cowichan\_\-mpi/cowichan\_\-mpi.cpp@{cowichan\_\-mpi/cowichan\_\-mpi.cpp}}
}
This file contains main method that drives the MPI implementation.

{\tt \#include \char`\"{}cowichan\_\-mpi.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__mpi}{cowichan\_\-mpi}
\begin{CompactList}\small\item\em Additional classes and functions specific to mpi implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\item 
bool \hyperlink{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}{cowichan\_\-mpi::get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end)
\item 
bool \hyperlink{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}{cowichan\_\-mpi::get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} rank)
\end{CompactItemize}

\subsection{Detailed Description}
This file contains main method that drives the MPI implementation.

\subsection{Function Documentation}
\hypertarget{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-mpi.cpp@{cowichan\_\-mpi.cpp}!main@{main}}
\index{main@{main}!cowichan_mpi.cpp@{cowichan\_\-mpi.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Main method - creates a \hyperlink{class_cowichan_m_p_i}{CowichanMPI} instance and executes \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{Cowichan::main}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \end{description}
\end{Desc}