### 2009-08-07 doc/latex/cowichan__openmp_2invperc_8cpp.tex

\hypertarget{cowichan__openmp_2invperc_8cpp}{
\section{cowichan\_\-openmp/invperc.cpp File Reference}
\label{cowichan__openmp_2invperc_8cpp}\index{cowichan\_\-openmp/invperc.cpp@{cowichan\_\-openmp/invperc.cpp}}
}
OpenMP (NOT parallel) invperc implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
{\tt \#include $<$queue$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__openmp_1_1_perc_point}{cowichan\_\-openmp::PercPoint}
\begin{CompactList}\small\item\em Percolation point. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
OpenMP (NOT parallel) invperc implementation.

The same as \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial::invperc}