### 2009-08-07 doc/latex/cowichan__openmp_2mandel_8cpp.tex

\hypertarget{cowichan__openmp_2mandel_8cpp}{
\section{cowichan\_\-openmp/mandel.cpp File Reference}
\label{cowichan__openmp_2mandel_8cpp}\index{cowichan\_\-openmp/mandel.cpp@{cowichan\_\-openmp/mandel.cpp}}
}
OpenMP mandel implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}{cowichan\_\-openmp::mandel\_\-calc} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} x, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} y)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP mandel implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{CowichanOpenMP::mandel} \end{Desc}