### 2009-08-07 doc/latex/cowichan__openmp_2norm_8cpp.tex

\hypertarget{cowichan__openmp_2norm_8cpp}{
\section{cowichan\_\-openmp/norm.cpp File Reference}
\label{cowichan__openmp_2norm_8cpp}\index{cowichan\_\-openmp/norm.cpp@{cowichan\_\-openmp/norm.cpp}}
}
OpenMP norm implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}{cowichan\_\-openmp::findMinMax} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{Point} $\ast$minPoint, \hyperlink{class_point}{Point} $\ast$maxPoint)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP norm implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{CowichanOpenMP::norm} \end{Desc}