### 2009-08-07 doc/latex/cowichan__openmp_2product_8cpp.tex

\hypertarget{cowichan__openmp_2product_8cpp}{
\section{cowichan\_\-openmp/product.cpp File Reference}
\label{cowichan__openmp_2product_8cpp}\index{cowichan\_\-openmp/product.cpp@{cowichan\_\-openmp/product.cpp}}
}
OpenMP product implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP product implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{CowichanOpenMP::product} \end{Desc}