### 2009-08-07 doc/latex/cowichan__openmp_2randmat_8cpp.tex

\hypertarget{cowichan__openmp_2randmat_8cpp}{
\section{cowichan\_\-openmp/randmat.cpp File Reference}
\label{cowichan__openmp_2randmat_8cpp}\index{cowichan\_\-openmp/randmat.cpp@{cowichan\_\-openmp/randmat.cpp}}
}
OpenMP randmat implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP randmat implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{CowichanOpenMP::randmat} \end{Desc}