### 2009-08-07 doc/latex/cowichan__openmp_2thresh_8cpp.tex

\hypertarget{cowichan__openmp_2thresh_8cpp}{
\section{cowichan\_\-openmp/thresh.cpp File Reference}
\label{cowichan__openmp_2thresh_8cpp}\index{cowichan\_\-openmp/thresh.cpp@{cowichan\_\-openmp/thresh.cpp}}
}
OpenMP thresh implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{CowichanOpenMP::thresh} \end{Desc}