### 2009-08-07 doc/latex/cowichan__openmp_2vecdiff_8cpp.tex

\hypertarget{cowichan__openmp_2vecdiff_8cpp}{
\section{cowichan\_\-openmp/vecdiff.cpp File Reference}
\label{cowichan__openmp_2vecdiff_8cpp}\index{cowichan\_\-openmp/vecdiff.cpp@{cowichan\_\-openmp/vecdiff.cpp}}
}
OpenMP vecdiff implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP vecdiff implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{CowichanOpenMP::vecdiff} \end{Desc}