### 2009-08-07 doc/latex/cowichan__openmp_2winnow_8cpp.tex

\hypertarget{cowichan__openmp_2winnow_8cpp}{
\section{cowichan\_\-openmp/winnow.cpp File Reference}
\label{cowichan__openmp_2winnow_8cpp}\index{cowichan\_\-openmp/winnow.cpp@{cowichan\_\-openmp/winnow.cpp}}
}
OpenMP winnow implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
{\tt \#include \char`\"{}sort.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}{cowichan\_\-openmp::mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$buckets)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP winnow implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{CowichanOpenMP::winnow} \end{Desc}