### 2009-08-07 doc/latex/cowichan__openmp_8cpp.tex

\hypertarget{cowichan__openmp_8cpp}{
\section{cowichan\_\-openmp/cowichan\_\-openmp.cpp File Reference}
\label{cowichan__openmp_8cpp}\index{cowichan\_\-openmp/cowichan\_\-openmp.cpp@{cowichan\_\-openmp/cowichan\_\-openmp.cpp}}
}
This file contains main method that drives the OpenMP implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__openmp_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\end{CompactItemize}

\subsection{Detailed Description}
This file contains main method that drives the OpenMP implementation.

\subsection{Function Documentation}
\hypertarget{cowichan__openmp_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-openmp.cpp@{cowichan\_\-openmp.cpp}!main@{main}}
\index{main@{main}!cowichan_openmp.cpp@{cowichan\_\-openmp.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__openmp_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Main method - creates a \hyperlink{class_cowichan_open_m_p}{CowichanOpenMP} instance and executes \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{Cowichan::main}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \end{description}
\end{Desc}