### 2009-08-07 doc/latex/cowichan__openmp_8hpp.tex

\hypertarget{cowichan__openmp_8hpp}{
\section{cowichan\_\-openmp/cowichan\_\-openmp.hpp File Reference}
\label{cowichan__openmp_8hpp}\index{cowichan\_\-openmp/cowichan\_\-openmp.hpp@{cowichan\_\-openmp/cowichan\_\-openmp.hpp}}
}
Open Multi-Processing (OpenMP) implementation of \hyperlink{class_cowichan}{Cowichan} problems.

{\tt \#include \char`\"{}../cowichan/cowichan.hpp\char`\"{}}\par
{\tt \#include $<$omp.h$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_cowichan_open_m_p}{CowichanOpenMP}
\begin{CompactList}\small\item\em Open Multi-Processing (OpenMP) implementation. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
Open Multi-Processing (OpenMP) implementation of \hyperlink{class_cowichan}{Cowichan} problems.