### 2009-08-07 doc/latex/cowichan__serial_2life_8cpp.tex

\hypertarget{cowichan__serial_2life_8cpp}{
\section{cowichan\_\-serial/life.cpp File Reference}
\label{cowichan__serial_2life_8cpp}\index{cowichan\_\-serial/life.cpp@{cowichan\_\-serial/life.cpp}}
}
Serial life implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__serial_93fd16bd1e6cb78235571ce617773639}{cowichan\_\-serial::sumNeighbours} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} first, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} r, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} c, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
Serial life implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{CowichanSerial::life} \end{Desc}