### 2009-08-07 doc/latex/cowichan__serial_2mandel_8cpp.tex

\hypertarget{cowichan__serial_2mandel_8cpp}{
\section{cowichan\_\-serial/mandel.cpp File Reference}
\label{cowichan__serial_2mandel_8cpp}\index{cowichan\_\-serial/mandel.cpp@{cowichan\_\-serial/mandel.cpp}}
}
Serial mandel implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{namespacecowichan__serial_4cde20338a4976564f4ceaf7a8a2c68e}{cowichan\_\-serial::mandel\_\-calc} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} x, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} y)
\end{CompactItemize}

\subsection{Detailed Description}
Serial mandel implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{CowichanSerial::mandel} \end{Desc}