### 2009-08-07 doc/latex/cowichan__serial_2product_8cpp.tex

\hypertarget{cowichan__serial_2product_8cpp}{
\section{cowichan\_\-serial/product.cpp File Reference}
\label{cowichan__serial_2product_8cpp}\index{cowichan\_\-serial/product.cpp@{cowichan\_\-serial/product.cpp}}
}
Serial product implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial product implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{CowichanSerial::product} \end{Desc}