### 2009-08-07 doc/latex/cowichan__serial_2randmat_8cpp.tex

\hypertarget{cowichan__serial_2randmat_8cpp}{
\section{cowichan\_\-serial/randmat.cpp File Reference}
\label{cowichan__serial_2randmat_8cpp}\index{cowichan\_\-serial/randmat.cpp@{cowichan\_\-serial/randmat.cpp}}
}
Serial randmat implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial randmat implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{CowichanSerial::randmat} \end{Desc}