### 2009-08-07 doc/latex/cowichan__serial_2thresh_8cpp.tex

\hypertarget{cowichan__serial_2thresh_8cpp}{
\section{cowichan\_\-serial/thresh.cpp File Reference}
\label{cowichan__serial_2thresh_8cpp}\index{cowichan\_\-serial/thresh.cpp@{cowichan\_\-serial/thresh.cpp}}
}
Serial thresh implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{CowichanSerial::thresh} \end{Desc}