### 2009-08-07 doc/latex/cowichan__serial_2vecdiff_8cpp.tex

\hypertarget{cowichan__serial_2vecdiff_8cpp}{
\section{cowichan\_\-serial/vecdiff.cpp File Reference}
\label{cowichan__serial_2vecdiff_8cpp}\index{cowichan\_\-serial/vecdiff.cpp@{cowichan\_\-serial/vecdiff.cpp}}
}
Serial vecdiff implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial vecdiff implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{CowichanSerial::vecdiff} \end{Desc}