### 2009-08-07 doc/latex/cowichan__serial_2winnow_8cpp.tex

\hypertarget{cowichan__serial_2winnow_8cpp}{
\section{cowichan\_\-serial/winnow.cpp File Reference}
\label{cowichan__serial_2winnow_8cpp}\index{cowichan\_\-serial/winnow.cpp@{cowichan\_\-serial/winnow.cpp}}
}
Serial winnow implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__serial_4a948387cf6c0310f0dc43d50fdef087}{cowichan\_\-serial::mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
Serial winnow implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{CowichanSerial::winnow} \end{Desc}