### 2009-08-07 doc/latex/cowichan__serial_8hpp.tex

\hypertarget{cowichan__serial_8hpp}{
\section{cowichan\_\-serial/cowichan\_\-serial.hpp File Reference}
\label{cowichan__serial_8hpp}\index{cowichan\_\-serial/cowichan\_\-serial.hpp@{cowichan\_\-serial/cowichan\_\-serial.hpp}}
}
Serial implementation of \hyperlink{class_cowichan}{Cowichan} problems.

{\tt \#include \char`\"{}../cowichan/cowichan.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_cowichan_serial}{CowichanSerial}
\begin{CompactList}\small\item\em Serial implementation. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
Serial implementation of \hyperlink{class_cowichan}{Cowichan} problems.