### 2009-08-07 doc/latex/cowichan__tbb_2gauss_8cpp.tex

\hypertarget{cowichan__tbb_2gauss_8cpp}{
\section{cowichan\_\-tbb/gauss.cpp File Reference}
\label{cowichan__tbb_2gauss_8cpp}\index{cowichan\_\-tbb/gauss.cpp@{cowichan\_\-tbb/gauss.cpp}}
}
TBB gauss implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_row_elimination}{cowichan\_\-tbb::RowElimination}
\begin{CompactList}\small\item\em Performs row elimination, i.e. eliminate i-th column in j-th row. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB gauss implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB::gauss} \end{Desc}