### 2009-08-07 doc/latex/cowichan__tbb_2hull_8cpp.tex

\hypertarget{cowichan__tbb_2hull_8cpp}{
\section{cowichan\_\-tbb/hull.cpp File Reference}
\label{cowichan__tbb_2hull_8cpp}\index{cowichan\_\-tbb/hull.cpp@{cowichan\_\-tbb/hull.cpp}}
}
TBB hull implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_maximum_distance}{cowichan\_\-tbb::MaximumDistance}
\begin{CompactList}\small\item\em Compute maximum distance for hull. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__tbb_08647459098335f4d3abda9054a99547}{cowichan\_\-tbb::quickhull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn)
\item 
void \hyperlink{namespacecowichan__tbb_689d689bc3d950d3dd7fb9df89d4156e}{cowichan\_\-tbb::hull\_\-split} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn, \hyperlink{class_point}{Point} $\ast$p1, \hyperlink{class_point}{Point} $\ast$p2)
\end{CompactItemize}

\subsection{Detailed Description}
TBB hull implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB::hull} \end{Desc}