### 2009-08-07 doc/latex/cowichan__tbb_2thresh_8cpp.tex

\hypertarget{cowichan__tbb_2thresh_8cpp}{
\section{cowichan\_\-tbb/thresh.cpp File Reference}
\label{cowichan__tbb_2thresh_8cpp}\index{cowichan\_\-tbb/thresh.cpp@{cowichan\_\-tbb/thresh.cpp}}
}
TBB thresh implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_max_reducer}{cowichan\_\-tbb::MaxReducer}
\begin{CompactList}\small\item\em Finds the maximum value in the image. \item\end{CompactList}\item 
class \hyperlink{classcowichan__tbb_1_1_histogram}{cowichan\_\-tbb::Histogram}
\begin{CompactList}\small\item\em This class calculates the histogram of a matrix. \item\end{CompactList}\item 
class \hyperlink{classcowichan__tbb_1_1_threshold}{cowichan\_\-tbb::Threshold}
\begin{CompactList}\small\item\em Creates a boolean array based on a cut-off value. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{CowichanTBB::thresh} \end{Desc}