### 2009-08-07 doc/latex/cowichan__tbb_2winnow_8cpp.tex

\hypertarget{cowichan__tbb_2winnow_8cpp}{
\section{cowichan\_\-tbb/winnow.cpp File Reference}
\label{cowichan__tbb_2winnow_8cpp}\index{cowichan\_\-tbb/winnow.cpp@{cowichan\_\-tbb/winnow.cpp}}
}
TBB winnow implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_point_count}{cowichan\_\-tbb::PointCount}
\begin{CompactList}\small\item\em Calculates the number of points to sort for winnow. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB winnow implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{CowichanTBB::winnow} \end{Desc}