### 2009-08-07 doc/latex/files.tex

\section{File List}
Here is a list of all documented files with brief descriptions:\begin{CompactList}
\item\contentsline{section}{cowichan/\hyperlink{cowichan_8cpp}{cowichan.cpp} (Implementation for \hyperlink{class_cowichan}{Cowichan} class and common routines for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan_8cpp}}{}
\item\contentsline{section}{cowichan/\hyperlink{cowichan_8hpp}{cowichan.hpp} (Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan_8hpp}}{}
\item\contentsline{section}{cowichan/\hyperlink{cowichan__defaults_8hpp}{cowichan\_\-defaults.hpp} (Defaults for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan__defaults_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_8cpp}{cowichan\_\-lt.cpp} (Entry point for the cowichan linuxtuples application )}{\pageref{cowichan__lt_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_8hpp}{cowichan\_\-lt.hpp} (Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs; LinuxTuples implementation )}{\pageref{cowichan__lt_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2gauss_8cpp}{gauss.cpp} (LinuxTuples gauss implementation )}{\pageref{cowichan__lt_2src_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{gauss_8hpp}{gauss.hpp} (LinuxTuples gauss header file )}{\pageref{gauss_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2half_8cpp}{half.cpp} (LinuxTuples half implementation )}{\pageref{cowichan__lt_2src_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{half_8hpp}{half.hpp} (LinuxTuples halving shuffle header file )}{\pageref{half_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2hull_8cpp}{hull.cpp} (LinuxTuples hull implementation )}{\pageref{cowichan__lt_2src_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{hull_8hpp}{hull.hpp} (LinuxTuples iterative convex hull header file )}{\pageref{hull_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2invperc_8cpp}{invperc.cpp} (LinuxTuples (actually serial) invperc implementation )}{\pageref{cowichan__lt_2src_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2life_8cpp}{life.cpp} (LinuxTuples life implementation )}{\pageref{cowichan__lt_2src_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{life_8hpp}{life.hpp} (LinuxTuples conway's game of life header file )}{\pageref{life_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2mandel_8cpp}{mandel.cpp} (LinuxTuples mandelbrot set implementation )}{\pageref{cowichan__lt_2src_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{mandel_8hpp}{mandel.hpp} (LinuxTuples mandelbrot set header file )}{\pageref{mandel_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2norm_8cpp}{norm.cpp} (LinuxTuples norm implementation )}{\pageref{cowichan__lt_2src_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{norm_8hpp}{norm.hpp} (LinuxTuples norm header file )}{\pageref{norm_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2outer_8cpp}{outer.cpp} (LinuxTuples outer implementation )}{\pageref{cowichan__lt_2src_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{outer_8hpp}{outer.hpp} (LinuxTuples outer product header file )}{\pageref{outer_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2product_8cpp}{product.cpp} (LinuxTuples product implementation )}{\pageref{cowichan__lt_2src_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{product_8hpp}{product.hpp} (LinuxTuples product header file )}{\pageref{product_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2randmat_8cpp}{randmat.cpp} (LinuxTuples randmat implementation )}{\pageref{cowichan__lt_2src_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{randmat_8hpp}{randmat.hpp} (LinuxTuples header file for the randmat cowichan problem )}{\pageref{randmat_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2sor_8cpp}{sor.cpp} (LinuxTuples successive over-relaxation implementation )}{\pageref{cowichan__lt_2src_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{sor_8hpp}{sor.hpp} (LinuxTuples successive over-relaxation header file )}{\pageref{sor_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2thresh_8cpp}{thresh.cpp} (LinuxTuples thresh implementation )}{\pageref{cowichan__lt_2src_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{thresh_8hpp}{thresh.hpp} (LinuxTuples histogram thresholding header file )}{\pageref{thresh_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{tuple__common_8cpp}{tuple\_\-common.cpp} (Implementation of common tuple application logic )}{\pageref{tuple__common_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{tuple__common_8hpp}{tuple\_\-common.hpp} (Header file for tuple applications )}{\pageref{tuple__common_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2vecdiff_8cpp}{vecdiff.cpp} (LinuxTuples vecdiff implementation )}{\pageref{cowichan__lt_2src_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{vecdiff_8hpp}{vecdiff.hpp} (LinuxTuples vector difference header file )}{\pageref{vecdiff_8hpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2winnow_8cpp}{winnow.cpp} (LinuxTuples winnow implementation )}{\pageref{cowichan__lt_2src_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{winnow_8hpp}{winnow.hpp} (LinuxTuples winnow header file )}{\pageref{winnow_8hpp}}{}
\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_8cpp}{cowichan\_\-mpi.cpp} (This file contains main method that drives the MPI implementation )}{\pageref{cowichan__mpi_8cpp}}{}
\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_8hpp}{cowichan\_\-mpi.hpp} (Boost Message Passing Interface (MPI) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__mpi_8hpp}}{}
\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2sort_8cpp}{sort.cpp} (Implementation of MPI sorting algorithms )}{\pageref{cowichan__mpi_2sort_8cpp}}{}
\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2sort_8hpp}{sort.hpp} (Sorting algorithms that use MPI )}{\pageref{cowichan__mpi_2sort_8hpp}}{}
\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2winnow_8cpp}{winnow.cpp} (MPI winnow implementation )}{\pageref{cowichan__mpi_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_8cpp}{cowichan\_\-openmp.cpp} (This file contains main method that drives the OpenMP implementation )}{\pageref{cowichan__openmp_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_8hpp}{cowichan\_\-openmp.hpp} (Open Multi-Processing (OpenMP) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__openmp_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2gauss_8cpp}{gauss.cpp} (OpenMP gauss implementation )}{\pageref{cowichan__openmp_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2half_8cpp}{half.cpp} (OpenMP half implementation )}{\pageref{cowichan__openmp_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2hull_8cpp}{hull.cpp} (OpenMP hull implementation )}{\pageref{cowichan__openmp_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2invperc_8cpp}{invperc.cpp} (OpenMP (NOT parallel) invperc implementation )}{\pageref{cowichan__openmp_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2life_8cpp}{life.cpp} (OpenMP life implementation )}{\pageref{cowichan__openmp_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2mandel_8cpp}{mandel.cpp} (OpenMP mandel implementation )}{\pageref{cowichan__openmp_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2norm_8cpp}{norm.cpp} (OpenMP norm implementation )}{\pageref{cowichan__openmp_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2outer_8cpp}{outer.cpp} (OpenMP outer implementation )}{\pageref{cowichan__openmp_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2product_8cpp}{product.cpp} (OpenMP product implementation )}{\pageref{cowichan__openmp_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2randmat_8cpp}{randmat.cpp} (OpenMP randmat implementation )}{\pageref{cowichan__openmp_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sor_8cpp}{sor.cpp} (OpenMP sor implementation )}{\pageref{cowichan__openmp_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sort_8cpp}{sort.cpp} (Implementation of OpenMP sorting algorithms )}{\pageref{cowichan__openmp_2sort_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sort_8hpp}{sort.hpp} (Sorting algorithms that use OpenMP )}{\pageref{cowichan__openmp_2sort_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2thresh_8cpp}{thresh.cpp} (OpenMP thresh implementation )}{\pageref{cowichan__openmp_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2vecdiff_8cpp}{vecdiff.cpp} (OpenMP vecdiff implementation )}{\pageref{cowichan__openmp_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2winnow_8cpp}{winnow.cpp} (OpenMP winnow implementation )}{\pageref{cowichan__openmp_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_8cpp}{cowichan\_\-serial.cpp} (This file contains main method that drives the serial implementation )}{\pageref{cowichan__serial_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_8hpp}{cowichan\_\-serial.hpp} (Serial implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__serial_8hpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2gauss_8cpp}{gauss.cpp} (Serial gauss implementation )}{\pageref{cowichan__serial_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2half_8cpp}{half.cpp} (Serial half implementation )}{\pageref{cowichan__serial_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2hull_8cpp}{hull.cpp} (Serial hull implementation )}{\pageref{cowichan__serial_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2invperc_8cpp}{invperc.cpp} (Serial invperc implementation )}{\pageref{cowichan__serial_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2life_8cpp}{life.cpp} (Serial life implementation )}{\pageref{cowichan__serial_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2mandel_8cpp}{mandel.cpp} (Serial mandel implementation )}{\pageref{cowichan__serial_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2norm_8cpp}{norm.cpp} (Serial norm implementation )}{\pageref{cowichan__serial_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2outer_8cpp}{outer.cpp} (Serial outer implementation )}{\pageref{cowichan__serial_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2product_8cpp}{product.cpp} (Serial product implementation )}{\pageref{cowichan__serial_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2randmat_8cpp}{randmat.cpp} (Serial randmat implementation )}{\pageref{cowichan__serial_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2sor_8cpp}{sor.cpp} (Serial sor implementation )}{\pageref{cowichan__serial_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2thresh_8cpp}{thresh.cpp} (Serial thresh implementation )}{\pageref{cowichan__serial_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2vecdiff_8cpp}{vecdiff.cpp} (Serial vecdiff implementation )}{\pageref{cowichan__serial_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2winnow_8cpp}{winnow.cpp} (Serial winnow implementation )}{\pageref{cowichan__serial_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_8cpp}{cowichan\_\-tbb.cpp} (This file contains main method that drives the TBB implementation )}{\pageref{cowichan__tbb_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_8hpp}{cowichan\_\-tbb.hpp} (Thread Building Blocks (TBB) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__tbb_8hpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2gauss_8cpp}{gauss.cpp} (TBB gauss implementation )}{\pageref{cowichan__tbb_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2half_8cpp}{half.cpp} (TBB half implementation )}{\pageref{cowichan__tbb_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2hull_8cpp}{hull.cpp} (TBB hull implementation )}{\pageref{cowichan__tbb_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2invperc_8cpp}{invperc.cpp} (TBB (NOT parallel) invperc implementation )}{\pageref{cowichan__tbb_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2life_8cpp}{life.cpp} (TBB life implementation )}{\pageref{cowichan__tbb_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2mandel_8cpp}{mandel.cpp} (TBB mandel implementation )}{\pageref{cowichan__tbb_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2norm_8cpp}{norm.cpp} (TBB norm implementation )}{\pageref{cowichan__tbb_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2outer_8cpp}{outer.cpp} (TBB outer implementation )}{\pageref{cowichan__tbb_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2product_8cpp}{product.cpp} (TBB product implementation )}{\pageref{cowichan__tbb_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2randmat_8cpp}{randmat.cpp} (TBB randmat implementation )}{\pageref{cowichan__tbb_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2sor_8cpp}{sor.cpp} (TBB sor implementation )}{\pageref{cowichan__tbb_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2thresh_8cpp}{thresh.cpp} (TBB thresh implementation )}{\pageref{cowichan__tbb_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2vecdiff_8cpp}{vecdiff.cpp} (TBB vecdiff implementation )}{\pageref{cowichan__tbb_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2winnow_8cpp}{winnow.cpp} (TBB winnow implementation )}{\pageref{cowichan__tbb_2winnow_8cpp}}{}
\end{CompactList}