### 2009-08-07 doc/latex/half_8hpp.tex

\hypertarget{half_8hpp}{
\section{cowichan\_\-lt/src/half.hpp File Reference}
\label{half_8hpp}\index{cowichan\_\-lt/src/half.hpp@{cowichan\_\-lt/src/half.hpp}}
}
LinuxTuples halving shuffle header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_half}{LTHalf}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples halving shuffle header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_f3bd58ad2653f06d242d473e7bffda0b}{CowichanLinuxTuples::half} \end{Desc}