### 2009-08-07 doc/latex/hierarchy.tex

\section{Class Hierarchy}
This inheritance list is sorted roughly, but not completely, alphabetically:\begin{CompactList}
\item \contentsline{section}{Cowichan}{\pageref{class_cowichan}}{}
\begin{CompactList}
\item \contentsline{section}{CowichanLinuxTuples}{\pageref{class_cowichan_linux_tuples}}{}
\item \contentsline{section}{CowichanMPI}{\pageref{class_cowichan_m_p_i}}{}
\item \contentsline{section}{CowichanOpenMP}{\pageref{class_cowichan_open_m_p}}{}
\item \contentsline{section}{CowichanSerial}{\pageref{class_cowichan_serial}}{}
\item \contentsline{section}{CowichanTBB}{\pageref{class_cowichan_t_b_b}}{}
\end{CompactList}
\item \contentsline{section}{cowichan\_\-tbb::GameOfLife}{\pageref{classcowichan__tbb_1_1_game_of_life}}{}
\item \contentsline{section}{cowichan\_\-tbb::Histogram}{\pageref{classcowichan__tbb_1_1_histogram}}{}
\item \contentsline{section}{cowichan\_\-tbb::MakeDominant}{\pageref{classcowichan__tbb_1_1_make_dominant}}{}
\item \contentsline{section}{cowichan\_\-tbb::Mandelbrot}{\pageref{classcowichan__tbb_1_1_mandelbrot}}{}
\item \contentsline{section}{cowichan\_\-tbb::MaximumDistance}{\pageref{classcowichan__tbb_1_1_maximum_distance}}{}
\item \contentsline{section}{cowichan\_\-tbb::MaxReducer}{\pageref{classcowichan__tbb_1_1_max_reducer}}{}
\item \contentsline{section}{cowichan\_\-tbb::MinMaxReducer}{\pageref{classcowichan__tbb_1_1_min_max_reducer}}{}
\item \contentsline{section}{cowichan\_\-tbb::Normalizer}{\pageref{classcowichan__tbb_1_1_normalizer}}{}
\item \contentsline{section}{cowichan\_\-tbb::NormDiff}{\pageref{classcowichan__tbb_1_1_norm_diff}}{}
\item \contentsline{section}{PercPoint}{\pageref{class_perc_point}}{}
\item \contentsline{section}{cowichan\_\-openmp::PercPoint}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
\item \contentsline{section}{cowichan\_\-tbb::PercPoint}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
\item \contentsline{section}{cowichan\_\-serial::PercPoint}{\pageref{classcowichan__serial_1_1_perc_point}}{}
\item \contentsline{section}{Point}{\pageref{class_point}}{}
\item \contentsline{section}{cowichan\_\-tbb::PointCount}{\pageref{classcowichan__tbb_1_1_point_count}}{}
\item \contentsline{section}{cowichan\_\-tbb::PointDistances}{\pageref{classcowichan__tbb_1_1_point_distances}}{}
\item \contentsline{section}{cowichan\_\-tbb::Product}{\pageref{classcowichan__tbb_1_1_product}}{}
\item \contentsline{section}{cowichan\_\-tbb::RandomGenerator}{\pageref{classcowichan__tbb_1_1_random_generator}}{}
\item \contentsline{section}{cowichan\_\-tbb::Relaxer}{\pageref{classcowichan__tbb_1_1_relaxer}}{}
\item \contentsline{section}{cowichan\_\-tbb::RowElimination}{\pageref{classcowichan__tbb_1_1_row_elimination}}{}
\item \contentsline{section}{cowichan\_\-tbb::Shuffle}{\pageref{classcowichan__tbb_1_1_shuffle}}{}
\item \contentsline{section}{cowichan\_\-tbb::Threshold}{\pageref{classcowichan__tbb_1_1_threshold}}{}
\item \contentsline{section}{TupleApplication}{\pageref{class_tuple_application}}{}
\begin{CompactList}
\item \contentsline{section}{LTBounds}{\pageref{class_l_t_bounds}}{}
\item \contentsline{section}{LTForward}{\pageref{class_l_t_forward}}{}
\item \contentsline{section}{LTFrequency}{\pageref{class_l_t_frequency}}{}
\item \contentsline{section}{LTHalf}{\pageref{class_l_t_half}}{}
\item \contentsline{section}{LTHull}{\pageref{class_l_t_hull}}{}
\item \contentsline{section}{LTLife}{\pageref{class_l_t_life}}{}
\item \contentsline{section}{LTMandel}{\pageref{class_l_t_mandel}}{}
\item \contentsline{section}{LTNorm}{\pageref{class_l_t_norm}}{}
\item \contentsline{section}{LTOuter}{\pageref{class_l_t_outer}}{}
\item \contentsline{section}{LTProduct}{\pageref{class_l_t_product}}{}
\item \contentsline{section}{LTRandmat}{\pageref{class_l_t_randmat}}{}
\item \contentsline{section}{LTSor}{\pageref{class_l_t_sor}}{}
\item \contentsline{section}{LTThresh}{\pageref{class_l_t_thresh}}{}
\item \contentsline{section}{LTVecdiff}{\pageref{class_l_t_vecdiff}}{}
\item \contentsline{section}{LTWinnow}{\pageref{class_l_t_winnow}}{}
\end{CompactList}
\item \contentsline{section}{WeightedPoint}{\pageref{class_weighted_point}}{}
\end{CompactList}