### 2009-08-07 doc/latex/hull_8hpp.tex

\hypertarget{hull_8hpp}{
\section{cowichan\_\-lt/src/hull.hpp File Reference}
\label{hull_8hpp}\index{cowichan\_\-lt/src/hull.hpp@{cowichan\_\-lt/src/hull.hpp}}
}
LinuxTuples iterative convex hull header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_hull}{LTHull}
\end{CompactItemize}
\subsection*{Defines}
\begin{CompactItemize}
\item 
\#define \hyperlink{hull_8hpp_ac10d34dd2d4faa7f96f7ed398c884f4}{NO\_\-POINT}~-1
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples iterative convex hull header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_501f87594d62af261ff0b1954a60ddc4}{CowichanLinuxTuples::hull} \end{Desc}

\subsection{Define Documentation}
\hypertarget{hull_8hpp_ac10d34dd2d4faa7f96f7ed398c884f4}{
\index{hull.hpp@{hull.hpp}!NO\_\-POINT@{NO\_\-POINT}}
\index{NO\_\-POINT@{NO\_\-POINT}!hull.hpp@{hull.hpp}}
\subsubsection[{NO\_\-POINT}]{\setlength{\rightskip}{0pt plus 5cm}\#define NO\_\-POINT~-1}}
\label{hull_8hpp_ac10d34dd2d4faa7f96f7ed398c884f4}

Indicator value to denote no valid point is stored.