### 2009-08-07 doc/latex/life_8hpp.tex

\hypertarget{life_8hpp}{
\section{cowichan\_\-lt/src/life.hpp File Reference}
\label{life_8hpp}\index{cowichan\_\-lt/src/life.hpp@{cowichan\_\-lt/src/life.hpp}}
}
LinuxTuples conway's game of life header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_life}{LTLife}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples conway's game of life header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_967138bbdead76eec50f31465dd9d5f1}{CowichanLinuxTuples::life} \end{Desc}