### 2009-08-07 doc/latex/mandel_8hpp.tex

\hypertarget{mandel_8hpp}{
\section{cowichan\_\-lt/src/mandel.hpp File Reference}
\label{mandel_8hpp}\index{cowichan\_\-lt/src/mandel.hpp@{cowichan\_\-lt/src/mandel.hpp}}
}
LinuxTuples mandelbrot set header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_mandel}{LTMandel}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples mandelbrot set header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_8774404842868381379ec627e8373cec}{CowichanLinuxTuples::mandel} \end{Desc}