### 2009-08-07 doc/latex/randmat_8hpp.tex

\hypertarget{randmat_8hpp}{
\section{cowichan\_\-lt/src/randmat.hpp File Reference}
\label{randmat_8hpp}\index{cowichan\_\-lt/src/randmat.hpp@{cowichan\_\-lt/src/randmat.hpp}}
}
LinuxTuples header file for the randmat cowichan problem.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_randmat}{LTRandmat}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples header file for the randmat cowichan problem.