### 2009-08-07 doc/latex/sor_8hpp.tex

\hypertarget{sor_8hpp}{
\section{cowichan\_\-lt/src/sor.hpp File Reference}
\label{sor_8hpp}\index{cowichan\_\-lt/src/sor.hpp@{cowichan\_\-lt/src/sor.hpp}}
}
LinuxTuples successive over-relaxation header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_sor}{LTSor}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples successive over-relaxation header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_aa2c469c3f520a5a56c5ec1ac9a5a6c6}{CowichanLinuxTuples::sor} \end{Desc}