### 2009-08-07 doc/latex/thresh_8hpp.tex

\hypertarget{thresh_8hpp}{
\section{cowichan\_\-lt/src/thresh.hpp File Reference}
\label{thresh_8hpp}\index{cowichan\_\-lt/src/thresh.hpp@{cowichan\_\-lt/src/thresh.hpp}}
}
LinuxTuples histogram thresholding header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_frequency}{LTFrequency}
\item 
class \hyperlink{class_l_t_thresh}{LTThresh}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples histogram thresholding header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_b8cf055ec3d42a1ccde6d818c5a43a74}{CowichanLinuxTuples::thresh} \end{Desc}