### 2009-08-07 doc/latex/vecdiff_8hpp.tex

\hypertarget{vecdiff_8hpp}{
\section{cowichan\_\-lt/src/vecdiff.hpp File Reference}
\label{vecdiff_8hpp}\index{cowichan\_\-lt/src/vecdiff.hpp@{cowichan\_\-lt/src/vecdiff.hpp}}
}
LinuxTuples vector difference header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_vecdiff}{LTVecdiff}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples vector difference header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_213185666ea7dd0c9abc59899b454086}{CowichanLinuxTuples::vecdiff} \end{Desc}