### 2009-08-11 doc/latex/annotated.tex
add comment to thresh, invperc small fix

\section{Class List}
Here are the classes, structs, unions and interfaces with brief descriptions:\begin{CompactList}
\item\contentsline{section}{\hyperlink{class_cowichan}{Cowichan} (Base class for all C++ implementations )}{\pageref{class_cowichan}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_linux_tuples}{CowichanLinuxTuples} }{\pageref{class_cowichan_linux_tuples}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_m_p_i}{CowichanMPI} (Boost Message Passing Interface (MPI) implementation )}{\pageref{class_cowichan_m_p_i}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_open_m_p}{CowichanOpenMP} (Open Multi-Processing (OpenMP) implementation )}{\pageref{class_cowichan_open_m_p}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_serial}{CowichanSerial} (Serial implementation )}{\pageref{class_cowichan_serial}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_t_b_b}{CowichanTBB} (Thread Building Blocks (TBB) implementation )}{\pageref{class_cowichan_t_b_b}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_game_of_life}{cowichan\_\-tbb::GameOfLife} (Ping-pong solution to game of life )}{\pageref{classcowichan__tbb_1_1_game_of_life}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_histogram}{cowichan\_\-tbb::Histogram} (This class calculates the histogram of a matrix )}{\pageref{classcowichan__tbb_1_1_histogram}}{}
\item\contentsline{section}{\hyperlink{class_l_t_bounds}{LTBounds} }{\pageref{class_l_t_bounds}}{}
\item\contentsline{section}{\hyperlink{class_l_t_forward}{LTForward} }{\pageref{class_l_t_forward}}{}
\item\contentsline{section}{\hyperlink{class_l_t_frequency}{LTFrequency} }{\pageref{class_l_t_frequency}}{}
\item\contentsline{section}{\hyperlink{class_l_t_half}{LTHalf} }{\pageref{class_l_t_half}}{}
\item\contentsline{section}{\hyperlink{class_l_t_hull}{LTHull} }{\pageref{class_l_t_hull}}{}
\item\contentsline{section}{\hyperlink{class_l_t_life}{LTLife} }{\pageref{class_l_t_life}}{}
\item\contentsline{section}{\hyperlink{class_l_t_mandel}{LTMandel} }{\pageref{class_l_t_mandel}}{}
\item\contentsline{section}{\hyperlink{class_l_t_norm}{LTNorm} }{\pageref{class_l_t_norm}}{}
\item\contentsline{section}{\hyperlink{class_l_t_outer}{LTOuter} }{\pageref{class_l_t_outer}}{}
\item\contentsline{section}{\hyperlink{class_l_t_product}{LTProduct} }{\pageref{class_l_t_product}}{}
\item\contentsline{section}{\hyperlink{class_l_t_randmat}{LTRandmat} }{\pageref{class_l_t_randmat}}{}
\item\contentsline{section}{\hyperlink{class_l_t_sor}{LTSor} }{\pageref{class_l_t_sor}}{}
\item\contentsline{section}{\hyperlink{class_l_t_thresh}{LTThresh} }{\pageref{class_l_t_thresh}}{}
\item\contentsline{section}{\hyperlink{class_l_t_vecdiff}{LTVecdiff} }{\pageref{class_l_t_vecdiff}}{}
\item\contentsline{section}{\hyperlink{class_l_t_winnow}{LTWinnow} }{\pageref{class_l_t_winnow}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_make_dominant}{cowichan\_\-tbb::MakeDominant} (Makes a given matrix diagonally dominant )}{\pageref{classcowichan__tbb_1_1_make_dominant}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_mandelbrot}{cowichan\_\-tbb::Mandelbrot} (Perform mandelbrot generation )}{\pageref{classcowichan__tbb_1_1_mandelbrot}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_maximum_distance}{cowichan\_\-tbb::MaximumDistance} (Compute maximum distance for hull )}{\pageref{classcowichan__tbb_1_1_maximum_distance}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_max_reducer}{cowichan\_\-tbb::MaxReducer} (Finds the maximum value in the image )}{\pageref{classcowichan__tbb_1_1_max_reducer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_min_max_reducer}{cowichan\_\-tbb::MinMaxReducer} (Performs the minimum and maximum reductions )}{\pageref{classcowichan__tbb_1_1_min_max_reducer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_normalizer}{cowichan\_\-tbb::Normalizer} (Normalizes points )}{\pageref{classcowichan__tbb_1_1_normalizer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_norm_diff}{cowichan\_\-tbb::NormDiff} (This class takes the 1-norm of the difference between two vectors )}{\pageref{classcowichan__tbb_1_1_norm_diff}}{}
\item\contentsline{section}{\hyperlink{class_perc_point}{PercPoint} }{\pageref{class_perc_point}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_perc_point}{cowichan\_\-tbb::PercPoint} (Percolation point )}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
\item\contentsline{section}{\hyperlink{classcowichan__serial_1_1_perc_point}{cowichan\_\-serial::PercPoint} (Percolation point )}{\pageref{classcowichan__serial_1_1_perc_point}}{}
\item\contentsline{section}{\hyperlink{classcowichan__openmp_1_1_perc_point}{cowichan\_\-openmp::PercPoint} (Percolation point )}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
\item\contentsline{section}{\hyperlink{class_point}{Point} (Two-dimensional point )}{\pageref{class_point}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_point_count}{cowichan\_\-tbb::PointCount} (Calculates the number of points to sort for winnow )}{\pageref{classcowichan__tbb_1_1_point_count}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_point_distances}{cowichan\_\-tbb::PointDistances} (Fills in matrix and vector with distances )}{\pageref{classcowichan__tbb_1_1_point_distances}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_product}{cowichan\_\-tbb::Product} (Performs matrix vector product )}{\pageref{classcowichan__tbb_1_1_product}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_random_generator}{cowichan\_\-tbb::RandomGenerator} (Random generator )}{\pageref{classcowichan__tbb_1_1_random_generator}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_relaxer}{cowichan\_\-tbb::Relaxer} (Performs one iteration of relaxation )}{\pageref{classcowichan__tbb_1_1_relaxer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_row_elimination}{cowichan\_\-tbb::RowElimination} (Performs row elimination, i.e. eliminate i-th column in j-th row )}{\pageref{classcowichan__tbb_1_1_row_elimination}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_shuffle}{cowichan\_\-tbb::Shuffle} (This class does a halving shuffle )}{\pageref{classcowichan__tbb_1_1_shuffle}}{}
\item\contentsline{section}{\hyperlink{structcowichan__mpi_1_1sum}{cowichan\_\-mpi::sum$<$ T $>$} (Calculates \hyperlink{structcowichan__mpi_1_1sum}{sum} of two elements )}{\pageref{structcowichan__mpi_1_1sum}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_threshold}{cowichan\_\-tbb::Threshold} (Creates a boolean array based on a cut-off value )}{\pageref{classcowichan__tbb_1_1_threshold}}{}
\item\contentsline{section}{\hyperlink{class_tuple_application}{TupleApplication} }{\pageref{class_tuple_application}}{}
\item\contentsline{section}{\hyperlink{class_weighted_point}{WeightedPoint} (Two-dimensional point with an integer weight )}{\pageref{class_weighted_point}}{}
\end{CompactList}