### 2009-08-28 paper/2009-09/compare.tex

\section{Comparing Parallel Programming Systems\label{s:questions}}

This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
It is presented here as an organizational aid;
when describing the PPS they have used,
implementors are encouraged to use the order and terminology of this list where appropriate.
We acknowledge that many of these questions will not apply to particular systems,
and encourage implementors to ask questions which we have left out.

\begin{enumerate}
\item History and Philosophy
  \begin{enumerate}
  \item For whom and for what purpose was the language designed?
  \end{enumerate}
\item Syntax
  \begin{enumerate}
  \item Model implemented (Figure~\ref{f:taxonomy})
  \item Class
    \begin{enumerate}
    \item library called from existing sequential language?
    \item syntactic extensions to existing sequential language?
    \item new language in Backus (Fortran/Algol/C) tradition?
    \item new (non-Backus) language?
    \end{enumerate}
  \item Concurrency mechanisms adhere to same philosophy as original
	language (where appropriate)?
  \item Readability?  Syntactic consistency?  Conciseness?
	(on an APL-to-Modula scale)
  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
  \end{enumerate}
\item Type Philosophy and Data Types
  \begin{enumerate}
  \item What primitive types does the language support?
  \item How are new types created?
  \item What intrinsic operations are defined on user-defined types?
  \item How strict is typing?
	How is type conversion done?
  \item Can type information be inspected at run-time?
  \item What concurrent operations on aggregates
	(such as lists and arrays) are intrinsic?
	What kinds of concurrent aggregate operations can users define?
  \item How are aggregates defined (e.g.\ by index set, by shape)?
	Can new kinds of aggregates (e.g.\ templates for arrays)
	be defined?
  \item How strict is typing during communication?
  \item What support is there for communication or manipulation
	of data of non-scalar or non-intrinsic types?
  \item Can structural information about aggregates
	(e.g.\ number of elements in a set, dimension of an array)
	be inspected at run-time?
  \item What built-in support is there for object classes and inheritance?
	For polymorphism?
	For class (as opposed to instance) data?
  \item Are functions first-class?
  \item Are continuations supported?  Are they first-class?
  \item Is dynamically-allocated memory supported?
	Is it done automatically?
	Is garbage-collection done automatically?
	Can a program ever create a dangling pointer?
	An alias?
  \end{enumerate}
\item Operators
  \begin{enumerate}
  \item What intrinsic operators are provided (brief list)?
	How strongly do these type-check their arguments?
  \item How are new operators defined?
  \item Are the precedence rules straightforward?
  \item Are side-effects/mutation allowed inside non-call expressions
	(e.g.\ C's side-effecting \verb`i++` construct)?
  \item Are their results platform-independent?
  \item Are the results of intrinsic parallel operations reproducible?
	Are they platform-independent?
  \item Are heterogeneous operations on aggregates allowed
	(e.g.\ every element of a list of polygons calculates its
	center of mass)?
	If so, how are these described, and how are they implemented?
  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
  \end{enumerate}
\item Control Flow
  \begin{enumerate}
  \item What synchronization is automatically imposed during execution
	(i.e.\ at what points can definite assertions about the state
	of the running program be made)?
  \item What control constructs does the language support?  Encourage?
  \item How is concurrency expressed?  How structured is this mechanism?
  \item Can the degree of concurrency be throttled (limited)?
	Is this done by the system, or by the user?
  \item How are exceptions handled/reported?
	Is it different for intrinsic operations and user-defined operations?
  \end{enumerate}
\item Subroutines, Scope, and Modularization
  \begin{enumerate}
  \item How efficient is subroutine call compared to in-line execution
	(i.e.\ how expensive are parameter passing and non-local control flow)?
  \item How much information about parameters is available (required)
	inside called subroutine
	(e.g.\ whether object is shared with other processes or private
	to executing process)?
  \item Are side-effecting procedures allowed?
  \item Can generic routines be written
	(e.g.\ to work on different types or array shapes)?
  \item Is there a restriction on return values (e.g.\ can a function return
	any object which could be declared)?
  \item Are there levels of scope?
	If multiple levels are permitted, how are references to stale scopes
	handled or prevented?
	Can information for declarations (such as size) be inherited (dynamically)
	from higher scopes?
  \item What synchronization requirements (if any) are imposed during calls
	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
	at the same time)?
  \item What support is there for modularization?
	For hiding implementation details?
	For sharing such information?
	For detecting and resolving name clashes?
  \end{enumerate}
\item Concurrent Programming Facilities
  \begin{enumerate}
  \item At what level(s) is the system concurrent?
    \begin{enumerate}
    \item implicit (the user shouldn't ask such things)
    \item operational (e.g.\ data-parallel addition)
    \item program counter (single executable, multiple states)
    \item procedural (heterogeneous control parallelism)
      \begin{enumerate}
      \item one execution thread per address space
      \item multiple threads per address space
      \end{enumerate}
    \end{enumerate}
  \item What mutual exclusion and synchronization primitives are provided?
  \item At what level(s) is data sharing visible to the programmer?
    \begin{enumerate}
    \item sequential consistency
    \item sequential consistency within processor clusters
    \item object-level consistency
    \item there is no sharing
    \end{enumerate}
  \item What is the conceptual granularity of operations?
    \begin{enumerate}
    \item synchronization after atomic instruction
    \item homogeneous operations with large-grain synchronization
    \item heterogeneous operations with large-grain synchronization
    \item completely asynchronous execution
    \end{enumerate}
  \item How do processes communicate data values?
	How do they synchronize?
	Are non-blocking operations (ones in which buffers might be
	in a visibly volatile state) allowed?
	How are they checked for termination?
  \item If data are communicated explicitly,
	how does the receiver choose what to accept?
    \begin{enumerate}
    \item by address (e.g.\ channel or port)
    \item by pattern matching on contents
    \item by explicit message tagging
    \item by provision of a typed buffer
    \item by provision of an untyped (byte-block) buffer
    \item in FIFO order
    \end{enumerate}
  \item How do users specify the number of processors on which to run?
	Must this be built into the program,
	or can it be decided at run-time?
	Is there support for executing on heterogeneous processors?
  \item Can the user specify the mapping of processes to processors?
	If so, how?
	How tightly coupled are the mapping of processes to processors,
	and the mapping of distributed data structures (if supported)?
  \item Can the user specify the mapping of data structures to processors?
        Can data structures be decomposed or partitioned?
  \item What support is there for operations on groups of processes, such as:
    \begin{enumerate}
    \item broadcast
    \item multicast (partial broadcast)
    \item barrier synchronization
    \end{enumerate}
  \item Can arbitrary process groups be created,
	or must they be structured in some way
	(e.g.\ as logical mesh)?
	Can operations involving disjoint groups be executed concurrently
	(e.g.\ a broadcast within one group,
	and a barrier within another)?
  \end{enumerate}
\item Input/Output
  \begin{enumerate}
  \item Is simple sequential (single-process) I/O simple to write?
  \item Is concurrent I/O supported?  If so, which flavours:
    \begin{enumerate}
    \item single-reader-broadcast/collect-single-writer
    \item homogeneously-sized or -strided segments
    \item heterogeneously-sized or -strided segments
    \item arbitrary independent operations
    \end{enumerate}
  \item What types can be read/written intrinsically?
	E.g.\ can an entire record (or array, or list)
	be read or written in a single operation?
  \end{enumerate}
\item Access to Routines in Other Languages and to the Hardware
  \begin{enumerate}
  \item Can subroutines written in other languages be called?
        Can subroutines in this language be called from other languages?
  \item Are there ways to access the run-time system directly,
	or similar low-level facilities?
  \item Is there high-level access to physical addresses and devices?
  \end{enumerate}
\item Practical Considerations
  \begin{enumerate}
  \item Are both batch and interactive processing supported?
  \item How reliable and efficient are existing implementations?
	How fast is it (i.e.\ how quickly does code compile and load)?
	Is it supported (who fixes it if it breaks)?
  \item Is there a standard for the system (from some standards organization)?
  \item Is the documentation adequate? For whom (CS graduates, computer-literate
	users, the general population)? Is training available?	Consulting?
  \item How much code has been written (lines of code or weeks of effort)?
        By whom (computer science undergraduates, professional engineers, etc.)?
  \item How long does it take the compiler to generate a ``small'' executable
        (such as {\tt{"hello, world"}})?
  \item How large is a ``small'' executable?
        How long does it take to load such an executable?
  \end{enumerate}
\item Supporting Tools
  \begin{enumerate}
  \item What is the development environment like?
        Can code be developed on a workstation and then moved to a parallel platform?
  \item Are there debugging aids, performance analysers, and management tools?
  \item How much help do these really provide?
  \end{enumerate}
\end{enumerate}