### 2009-08-28 paper/2009-09/concl.tex

\section{Conclusion\label{s:conclude}}

A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
\begin{center}
{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
\end{center}
This release includes two parallel versions,
both based on the POSIX threads package.
The first uses repeated fork-and-join parallelism,
while the second creates a fixed number of threads at the beginning of the program,
then synchronizes them repeatedly using barriers.

We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
and, if there is sufficient interest, to publish this material in book form.
If you would like to contribute to this effort, please contact either of the first two authors.