### 2009-08-28 paper/2009-09/intro.tex

\section{Introduction\label{s:intro}}

Twenty years ago,
Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
A similar survey today could well uncover twice as many,
but a poll of application programmers would probably find that
only a handful are in general use.

The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
Part of the blame undoubtedly lies with the diversity of hardware,
and the attendant lack of portability,
alluded to above.
A second reason is that an academic researcher's principal interest is developing new languages,
not supporting existing ones.
By the time a system is 90\% complete,
its author is already thinking about its successor.
Few application programmers are willing to invest effort porting codes
to systems which are not properly supported,
and liable to change or evaporate without notice.
Third,
many research groups have chosen to explore the potential of parallelism
without consideration for compatibility with older languages.
Application programmers must therefore face the prospect of completely re-writing their programs,
rather than simply porting them.
As the next section argues,
this can obviate any performance that might be gained from parallelization.
Finally,
many PPSs simply cannot support the software engineering load of large applications.
Many provide only a single global namespace,
do not include I/O capabilities,
require users to marshal and unmarshal data explicitly,
or do not permit hardware to be timeshared.

The net result is that many PPSs are never used for anything more complicated than
generating the Mandelbrot Set,
simulating the Game of Life,
multiplying dense matrices,
or finding all solutions to the $N$-queens problem for some small $N$.
Lacking the feedback that comes from more demanding use,
few systems are improvements over their predecessors
in the way that successive versions of operating systems have improved on one another,
or the languages in the Algol$\rightarrow$Modula-$n$ family
have become both more powerful and more secure.
Most applications programmers therefore choose to play it safe
and stick to parallel programming's lowest common denominators:
data-parallel Fortran,
and message-passing libraries.

The historical reasons for tolerating this situation are becoming less valid with each passing year.
When only a few parallel computers existed,
and these were of widely different types,
software standardization simply didn't make sense.
Similarly,
when supercomputers were very expensive,
the cost of software development seemed relatively smaller,
and the requirement for maximum performance relatively greater,
than in sequential computing.
As \cite{b:social-limits-speed} points out,
however, these arguments are losing force.
Now,
as in sequential computing,
the first question is not ``How do I write programs for this machine?'',
but ``Will this machine run my programs?''
As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.

But herein lies a danger.
History shows that it is extremely difficult
to displace a programming language or system once it becomes established.
Cobol and MS-DOS are the most frequently cited examples,
but many systems which should have been superceded long ago have managed
not only to cling to life,
but to grow in influence.
(The fact that the mail signature
``The last good thing written in C was Schubert's Ninth Symphony''
can elicit smiles simply proves that even the truth can be funny.)
Established systems evolve slowly, if at all.
It would therefore seem that the best time to guarantee ourselves
a good parallel programming paradigm is right now,
before a large number of large applications are created and take on lives of their own.

We therefore propose to use a set of simple programming problems to compare
the usability of parallel programming systems.
Competent programmers, fluent in a specific system,
will implement solutions to these problems and report their experiences
in terms of development time, code size and clarity, and runtime efficiency.
These problems are on the scale of the toy applications described several paragraphs ago
both to keep implementation time low
and because we feel that any PPS produced today ought to be able to support such simple programs.
As the second author put it,
``We're all sick of seeing these,
but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
One significant innovation in this work is that
we also require implementations of our suite to chain problems together,
so that the output of one problem is the input to another.
This will allow us to assess how well each PPS supports modularization and code re-use
(or, more realistically, how much extra effort programmers must invest
to achieve these goals when using a particular PPS).

Section~\ref{s:corollary} motivates this work by showing
how poor usability can limit the realizable parallelism in a program.
Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
Section~\ref{s:toys} then presents the suite of problems we intend to use.
In recognition of the original work done by Feo and others on
the Salishan Problems~\cite{b:salishan}, this suite is called
the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
	Like Salishan, the word is a Northwest Coast Indian place name.}.

\subsection*{History and Acknowledgments}

The first author originally developed this suite in the mid-1990s,
inspired both by \cite{b:salishan} and by
discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
Henri Bal (Vrije Universiteit, Amsterdam),
and others.
Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.