### 2009-08-28 paper/2009-09/issues.tex

\section{Other Issues\label{s:issues}}

As described in Section~\ref{s:method-softeng},
implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
This section describes the flow of control and data between these chained problems,
and discusses several other important aspects of this suite,
including I/O and visualization.

\subsection{Chained Problem Sequences\label{s:issues-chain}}

The chained implementation of these problems
executes individual toys in the following order:
\begin{enumerate}
\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
	\label{e:chain-choice-1}
\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
	Like $I$, $B$ has $r$ rows and $c$ columns.
	\label{e:chain-choice-2}
\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
	This step overwrites the Boolean matrix $B$.
\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
	as described in Section~\ref{s:toys-outer}.
\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
	In a parallel system, these two problems should if possible be executed concurrently.
\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
	In a parallel system, these two problems should if possible be executed concurrently.
\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
	This measures the agreement between the solutions found by the two methods.
\end{enumerate}
Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
there are four possible chained sequences.
A minimal implementation will provide one of these,
while a full implementation will be a single program which can execute any of the four sequences.
Intermediate implementations providing all four sequences as separate programs are also acceptable.

\subsection{Input and Output\label{s:issues-io}}

I/O is an important part of programming,
but is often treated as being of secondary importance by language designers.
This suite requires all stand-alone toys to read input values from files, and write results back;
the chained version must be able to checkpoint intermediate results between toys.
Finally,
implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.

The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
A file containing a vector begins with 1 positive integer,
which specifies the number of elements in the vector;
a file containing a matrix begins with 2 positive integers,
which specify the number of rows and columns in the matrix respectively.
Elements of the vector or matrix then appear one per line,
from least-indexed to greatest-indexed
(and in row-major order for matrices).

The formats described above are required to be human-readable.
Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
using whatever file formats are convenient.
This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
which most probably be used for checkpointing intermediate results in real programs.

\begin{figure}
% \epsfxsize=0.75\textwidth
% \begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
\caption{Cowichan File Formats\label{f:io-formats}}
\end{figure}

\subsection{Visualization\label{s:issues-vis}}

Visualization is an increasingly important part of scientific and numerical computing.
Including it as an optional part of this problem set is also a good way to exercise both
the use of libraries within a parallel programming system,
and the ease with which a programming system can support variant versions of a program.

Our requirements for the graphics interface are:
\begin{enumerate}
\item	Visualization for individual toys should be as simple as possible,
	and should require as few changes to the basic toy code as possible.
\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
	The latter should create all images in a single window.
\item	A single source file should implement both the graphical and non-graphical versions of each toy.
\end{enumerate}
The first two points require the system to distinguish between the actual drawing code
(e.g.\ drawing a vector of points as circles)
and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
While the stand-alone toys use a whole window for drawing,
the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
At present,
this is implemented by using a call counter
in the interfaces for the individual toys and in the chained module.
The actual drawing routines called from within each toy could then be the same.

The present visualization routines were originally written by Anil Sukul,
a student at the Vrije Universiteit, Amsterdam.
The package uses the public domain graphics library VOGLE,
available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
as well as screen control routines (such as double buffering).
VOGLE was chosen because of its simplicity,
and because it includes Pascal and Fortran interfaces.

\subsection{Reproducibility\label{s:issues-reproduce}}

Reproducibility is an important issue for parallel programming systems.
While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
it can also reduce the expressiveness or performance of the system.

In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
or the order in which those processors carry out particular operations.

Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
If several cells have this value, implementations may choose one arbitrarily.
Thus, different implementations may produce very different fractal shapes.
In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
although the values in the resulting matrix then depend on the number of processors used.