### 2009-08-28 paper/2009-09/paper.tex

\documentstyle[epsf,11pt]{article}

\begin{document}

\input{titlepage}

\vspace{\baselineskip}

\input{abstract}
\input{intro}
\input{quant}
\input{method}
\input{problems}
\input{issues}
\input{concl}

\bibliographystyle{plain}
\newcommand{\bibea}{{\em{et~al.}}}
\bibliography{paper}

\appendix

\input{cliches}
\input{compare}

\end{document}