### 2009-08-28 paper/2009-09/prob-convex.tex

\subsection{{\tt{convex}}:
	Convex Hull
	\label{s:toys-convex}}

This module takes a list of two-dimensional points and reorders them by
doing multiple convex hull computations. The convex hull is the boundary of
the minimal convex set containing a given non-empty finite set of points
in the plane. In other words, all points not in the convex hull are
enclosed in the convex hull polygon. At each step the convex hull points
are taken out of the input list and are put into the output list. The
computation terminates when there are no more points left in the input
list.  Inputs are:

\begin{description}
\item[{\tt{pointsIn}}:]
	the vector of input points.
\end{description}
Its output is:
\begin{description}
\item[{\tt{pointsOut}}:]
	the vector of output points (a permutation of the input).
\end{description}