### 2009-08-28 paper/2009-09/prob-gauss.tex

\subsection{{\tt{gauss}}:
	Gaussian Elimination
	\label{s:toys-gauss}}

This module solves a matrix equation $AX=V$ for
a dense, symmetric, diagonally dominant matrix $A$
and an arbitrary vector non-zero $V$
using explicit reduction.
(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
there is a well-formed solution to the equation.)
Inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the real matrix $A$.
\item[{\tt{target}}:]
	the real vector $V$.
\item[{\tt{nelts}}:]
	the number of values in each vector, and the size of the matrix along each axis.
\end{description}
Its output is:
\begin{description}
\item[{\tt{solution}}:]
	a real vector containing the solution $X$.
\end{description}