### 2009-08-28 paper/2009-09/prob-half.tex

\subsection{{\tt{half}}:
	Two-Dimensional Shuffle
	\label{s:toys-half}}

This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
shuffles them,
and then repeats this operation along the other axis.
Values in odd-numbered locations are collected at the low end of each row or column,
while values in even-numbered locations are moved to the high end.
An example transformation is:

\begin{eqnarray*}
\begin{array}{cccc}
a & b & c & d \\
e & f & g & h \\
i & j & k & l
\end{array}
& = &
\begin{array}{cccc}
a & c & b & d \\
i & k & j & l \\
e & g & f & h
\end{array}
\end{eqnarray*}

Note that how an array element is moved depends only on whether its location index is odd or even,
not on whether its value is odd or even.

The inputs to this module are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix containing shuffled values.
\end{description}