### 2009-08-28 paper/2009-09/prob-norm.tex

\subsection{{\tt{norm}}:
	Point Location Normalization
	\label{s:toys-norm}}

This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
Its inputs are:
\begin{description}
\item[{\tt{points}}:]
	a vector of point locations.
\item[{\tt{nelts}}:]
	the number of elements in the point vector.
\end{description}
Its output is:
\begin{description}
\item[{\tt{points}}:]
	a vector of normalized point locations.
\end{description}
If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
then the normalization equation is:
\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
$y$ coordinates are normalized in the same fashion.