### 2009-08-28 paper/2009-09/prob-outer.tex

\subsection{{\tt{outer}}:
	Outer Product
	\label{s:toys-outer}}

This module turns a vector containing point positions
into a dense, symmetric, diagonally dominant matrix
by calculating the distances between each pair of points.
It also constructs a real vector whose values are
the distance of each point from the origin.
Inputs are:
\begin{description}
\item[{\tt{points}}:]
	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
\item[{\tt{nelts}}:]
	the number of points in the vector, and the size of the matrix along each axis.
\end{description}
Its outputs are:
\begin{description}
\item[{\tt{matrix}}:]
	a real matrix, whose values are filled with inter-point distances.
\item[{\tt{vector}}:]
	a real vector, whose values are filled with origin-to-point distances.
\end{description}
Each matrix element $M_{i,j}$ such that $i \neq j$
is given the value $d_{i, j}$,
the Euclidean distance between point $i$ and point $j$.
The diagonal values $M_{i, i}$ are then set to
{\tt{nelts}} times the maximum off-diagonal value
to ensure that the matrix is diagonally dominant.
The value of the vector element $v_i$ is set to
the distance of point $i$ from the origin,
which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.