### 2009-08-28 paper/2009-09/prob-product.tex

\subsection{{\tt{product}}:
	Matrix-Vector Product
	\label{s:toys-product}}

Given a matrix $A$,
a vector $V$,
and an assumed solution $X$ to the equation $AX=V$,
this module calculates the actual product $AX={V^{\prime}}$,
and then finds the magnitude of the error.
Inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the real matrix $A$.
\item[{\tt{actual}}:]
	the real vector $V$.
\item[{\tt{candidate}}:]
	a real vector containing the supposed solution.
\item[{\tt{nelts}}:]
	the number of values in each vector, and the size of the matrix along each axis.
\end{description}
The output of this function is:
\begin{description}
\item[{\tt{e}}:]
	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
\end{description}