### 2009-08-28 paper/2009-09/prob-randmat.tex

\subsection{{\tt{randmat}}:
	Random Number Generation
	\label{s:toys-rng}}

This module fills a matrix with pseudo-random integers.
The inputs to this module are:
\begin{description}
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{s}}:]
	the random number generation seed.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix filled with random values.
\end{description}
Note that,
as in all problems,
the output is required to be independent of the number of processors used.
Generating new seed values from the given seed,
and running one copy of the random number generator
on each processor,
is therefore unlikely to qualify as a solution.