### 2009-08-28 paper/2009-09/prob-thresh.tex

\subsection{{\tt{thresh}}:
	Histogram Thresholding
	\label{s:toys-thresh}}

This module performs histogram thresholding on an image.
Given an integer image $I$ and a target percentage $p$,
it constructs a binary image $B$
such that $B_{i,j}$ is set
if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
The general idea is that an image's histogram should have 2 peaks,
one centered around the average foreground intensity,
and one centered around the average background intensity.
This program attempts to set
a threshold between the two peaks in the histogram
and select the pixels above the threshold.
This module's inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the integer matrix to be thresholded.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix and mask.
\item[{\tt{percent}}:]
	the minimum percentage of cells to retain.
\end{description}
Its output is:
\begin{description}
\item[{\tt{mask}}:]
	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
\end{description}