### 2009-08-28 paper/2009-09/prob-winnow.tex

\subsection{{\tt{winnow}}:
	Weighted Point Selection
	\label{s:toys-winnow}}

This module converts a matrix of integer values
to a vector of points,
represented as $x$ and $y$ coordinates.
Its inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix, whose values are used as masses.
\item[{\tt{mask}}:]
	a Boolean matrix showing which points are eligible for consideration.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{nelts}}:]
	the number of points to select.
\end{description}
Its output is:
\begin{description}
\item[{\tt{points}}:]
	a vector of $(x,y)$ points.
\end{description}
Each location where {\tt{mask}} is {\tt{true}}
becomes a candidate point,
with a weight equal to the integer value in {\tt{matrix}} at that location
and $x$ and $y$ coordinates
equal to its row and column indices.
These candidate points are then sorted into increasing order by weight,
and {\tt{nelts}} evenly-spaced points selected to create the result vector.