### 2009-08-28 paper/2009-09/problems.tex

\section{The Cowichan Problems\label{s:toys}}

This section describes the problems making up the full problem suite.
As discussed above,
the implementation of each is to be used both as a stand-alone program
and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.

\begin{table}
\begin{center}
\begin{tabular}{cl|cl}
Code	& Toy				& Code	& Toy				\\
\hline
C	& convex hull			& O	& outer product			\\
G	& Gaussian elimination		& P	& matrix-vector product		\\
H	& halving shuffle		& R	& random matrix generation	\\
I	& invasion percolation		& S	& successive over-relaxation	\\
L	& Game of Life			& T	& image thresholding		\\
M	& Mandelbrot Set		& V	& vector difference		\\
N	& point normalization		& W	& point value winnowing
\end{tabular}
\caption{Single-Character Codes for Toys\label{t:toy-codes}}
\end{center}
\end{table}

\input{prob-convex}
\input{prob-gauss}
\input{prob-half}
\input{prob-invperc}
\input{prob-life}
\input{prob-mandel}
\input{prob-norm}
\input{prob-outer}
\input{prob-product}
\input{prob-randmat}
\input{prob-sor}
\input{prob-thresh}
\input{prob-vecdiff}
\input{prob-winnow}