### 2009-08-31 paper/2009-09/method.tex

\section{Methodology\label{s:method}}

Our aim is to assess and compare the usability of parallel programming systems.
Here,
we discuss issues related to choosing the problems that make up our benchmark suite,
how to assess particular implementations,
and other purposes which implementations of these might serve.
Recognizing that our problems are extremely simple,
we refer to them as ``toys''.

\subsection{Criteria for Selection\label{s:method-criteria}}

Our criteria for including toys are:

\begin{enumerate}

\item	Each toy should require no more than an hour
	to write and test in a well-supported sequential language.
	We feel that making individual problems more complicated will only discourage uptake.

\item	Correctness must be easy to verify.
	Toys whose output is easily visualized are therefore preferred,
	as are toys whose results are insensitive to floating-point effects.

\item	Speedup must similarly be easy to measure---while machine performance is not our focus,
	we are generally uninterested in parallel implementations that are slower than
	the sequential originals.

\item	At least some toys should not be ``infinitely scalable''.
	Many real-world applications are not,
	and this suite should reflect such limitations.

\item	At least some toys should require I/O,
	since this important aspect of real-world programming is often neglected by PPS designers.

\item	There should be some overlap in the toys' implementations,
	so that implementors can demonstrate how well
	their chosen systems take advantage of opportunities for code re-use.

\item	Together,
	the toys in the suite must exercise a wide range of parallel clich\'{e}s
	(Appendix~\ref{s:ops} and ~\ref{s:memory}).
	In particular,

\item	Toys should be specified by inputs and outputs rather than algorithmically,
	i.e.,
	``sort $N$ integers'' rather than ``parallelize quicksort'',
	so that implementors can choose algorithms that are ``natural'' for their systems.
	Implementations that parallelize a grossly inefficient sequential algorithm
	should expect to be criticized for doing so;
	see Section~\ref{s:toys-invperc} for more discussion of this.

\end{enumerate}

\subsection{Software Engineering Issues\label{s:method-softeng}}

The ``single algorithm per program'' model of many benchmarks is not representative of real programs,
which often contain several qualitatively different phases or modules.
A full implementation of this suite will therefore have two parts.
In the first,
each toy will be implemented as a stand-alone program.
In the second,
toys will be chained together as shown in Figure~\ref{f:chaining}.
This will test the ease with which heterogeneous parallelism can be mixed within a single program.
It will also show how well the system supports code re-use and information hiding,
which are crucial to the development of large programs.

\begin{figure}
% \epsfxsize=12cm
% \begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
\caption{Chaining\label{f:chaining}}
\end{figure}

Where possible,
chaining should execute toys concurrently.
Some parallel programming systems impose extraneous constraints on programs,
e.g., require all processes to participate in every barrier synchronization,
or require the same executable to be loaded onto each processor.
These constraints can limit the exploitation of potential concurrency.
Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.

\subsection{Sizing\label{s:method-size}}

One crucial aspect of the specification of toys is
the way in which the sizes of problems are determined.
In a \emph{frozen} model,
the actual size of each problem and/or the number of processors available would be compiled into each toy.
A fully \emph{fluid} implementation, by contrast, would allow these sizes to be specified at run-time.
If a system does not support fully-fluid implementations,
it must at least use an intermediate model (which might be called \emph{slushy}),
in which the maximum size of individual problems is specified during compilation,
but the actual size of a problem is only determined when the toy begins to execute.

\subsection{Assessing Complexity\label{s:method-complex}}

There are at least two different ways in which usability could be compared.
One would be to measure the performance achieved by an ``average'' programmer
as a function of time on each toy.
By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
we could in principle find the value of $p_{1/2}$---the programming time required to achieve
half of a machine's peak performance\footnote{Note that
	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
	it is unlikely that the halfway mark would ever be reached.}
for a particular combination of programming system and problem type.
However,
the number of implementors is likely to be too small for averages to be meaningful.

Alternatively,
we could use code metrics
to compare implementations of various toys.
This is also dubious,
as \cite{b:el-emam-metrics} has shown that code metrics do little more than measure lines of code.

We therefore intend to compare implementations through group consensus,
e.g.,
through interviews with implementors
and by asking other programmers who are familiar with the base languages used
to read and comment on the parallel versions of the toys.
We will also compare the length of code in parallel and sequential implementations,
though we realize that this can exaggerate the impact of parallelization.

\subsection{Other Uses for Implementations\label{s:method-uses}}

This suite is intended for assessing the usability of parallel programming systems,
but we envisage other uses.
First, this suite will indicate
what a mature parallel programming system should be able to support.
In particular,
we will ask implementors to describe how they debugged correctness and performance problems,
and in particular what tools they used.
(The lack of useful debugging tools is a chronic complaint among parallel programmers.)

This suite should also be useful for education,
since toys will be small enough to be understood quickly,
and parallelizing them should be a suitable classroom exercise
in a senior undergraduate course on parallel computing.

\subsection{Some Criticisms\label{s:method-criticism}}

\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}

Implementations of benchmark suites have usually been coded for absolute speed.
They are not representative of ``normal'' practice,
in which reasonable performance is the usual goal,
and so are inappropriate for this work.
In addition,
most existing benchmark suites are too large to be implemented or ported quickly
(e.g.\ SPEC~\cite{b:bench-over}),
over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
or concerned with a limited range of applications.
Finally,
most exist only in C or Fortran (or both);
many interesting parallel programming systems are built on top of other languages.
We do not want the effort required to translate several large programs
to discourage researchers from participating in this exercise.

\noindent {\em{Won't the results depend primarily on programmer ability?}\/}

Yes,
but no more than the results from performance benchmarks.
A longer answer is that
since we intend to measure the combination of code complexity and performance achieved,
rather than performance as a function of development time,
programmers will be allowed to revise and improve their implementations when and as desired.
We hope this will encourage the evolutionary improvement seen in some online programming contests,
e.g., \cite{b:gulley-contest}.