### 2009-09-01 paper/2009-09/paper.tex
Tidying up problem descriptions

\documentstyle[epsf,11pt]{article}

\begin{document}

\newcommand{\bibea}{{\em{et~al.}}}
\newcommand{\inputspec}{\noindent \bf{Inputs:}}
\newcommand{\outputspec}{\noindent \bf{Outputs:}}

\input{titlepage}

\vspace{\baselineskip}

\input{abstract}
\input{intro}
\input{quant}
\input{method}
\input{problems}
\input{issues}
\input{concl}

\bibliographystyle{plain}
\bibliography{paper}

\appendix

\input{ops}
\input{memory}

\end{document}